---
title:  "CA-WRF-regional-runs"
author: "Xiulin Gao"
email:  "xiulingao@lbl.gov"
date:   "10-Nov-2022"
---
## Introduction
This script is used to process regional runs in California grasslands. 
We want to benchmark model simulations of annual burned fraction, GPP, and LAI in annual grasslands. 
This scripts uses a set of scripts developed by Marcos Longo at https://github.com/mpaiao/FATES_Utils;
part of this script is also adapted from Marcos Longo's post-processing script, changes are
made to do post-processing specifically for 1PFT regional simulations. 



```{r, label = 'reset-R',message=FALSE, results='hide'}
# Unload all packages except for the R default ones
plist = names(sessionInfo()$otherPkgs)
if (length(plist) > 0){
   dummy = sapply(X=paste0("package:",plist),FUN=detach,character.only=TRUE,unload=TRUE)
}#end if (length(plist) > 0)


# Remove all variables
rm(list=ls())

# Reset warnings
options(warn=0)

# Close all plots
invisible(graphics.off())

# Clean up
invisible(gc())
```






```{r, label="path-setting"}

home_path   = path.expand("~")
main_path   = file.path("/Volumes/easystore/derecho/oak-grass/trans/mort/")
util_path   = file.path(home_path,"Util/RUtils")


cases       = c( #"decidoak-project-N01T02_CLM_FATES.clm2.nc",
                 #"default-params-struc.nc"
                "default-params-trans_CLM_FATES.clm2.nc"
              )

case_files   = file.path(main_path, cases)
ncase        = length(cases)
case_names   = c(1L)
case_desc    = c( "Tonzi-params-trans")
names(case_desc) = c("1")
file_desc = "default-trans"

#Observation path
obs_main    = file.path("~/Google Drive/My Drive/CA-grassland-simulationDoc/benchmark")
site_main   = file.path("~/Google Drive/My Drive/CA-OakWoodland/benchmark")
frap_main   = file.path(obs_main,"FRAP")
ras_main    = file.path(obs_main,"CA-GPP-LAI-005degree")

site_file   = file.path(site_main,"tonzi_2001-2022.csv")
lai_path    = file.path(ras_main,"LAI-onWRFgrids.csv")
gpp_path    = file.path(ras_main,"GPP-onWRFgrids.csv")
frap_file   = file.path(frap_main,"Bfrac-resampled-OnWRFDomain.csv")
short_file  = file.path(frap_main,"BArea-resampled-OnWRFDomain_SHORT.csv")
mass_path   = file.path(site_main,"WRF_CA-HarwoodBiomass.csv")
bsarea_path = file.path(site_main,"WRF_oakBasalArea.csv")
cancov_path = file.path(site_main,"WRF_CA-HarwoodCanopyCover.csv")
wrf_path    = file.path("~/Google Drive/My Drive/9km-WRF-1980-2020/1981-01.nc")
clim_main   = file.path(obs_main,"global-historical-climate/wc2.1_10m_bio")

```






```{r,label="make-dir"}

#plot path
plot_main  = file.path(main_path,"Figures")
# Output path for time series
tsage_path    = file.path(plot_main,"tseries_age"         )
tsdbh_path    = file.path(plot_main,"tseries_dbh"         )
tspft_path    = file.path(plot_main,"tseries_pft"         )
tsmort_path   = file.path(plot_main,"tseries_mort"        )
tsnppo_path   = file.path(plot_main,"tseries_npp_organ"   )
tsauto_path   = file.path(plot_main,"tseries_auto_resp"   )
tsdap_path    = file.path(plot_main,"ts_heat_dbh+pft"     )
tsdphen_path  = file.path(plot_main,"ts_drought_phenology")
tssoil_path   = file.path(plot_main,"ts_heat_soil"        )
tstheme_path  = file.path(plot_main,"tseries_theme"       )
secy_path     = file.path(plot_main,"seasonal-cycle"      )
sens_path     = file.path(plot_main,"variance-decomp"     )

# Create paths for time series
dummy = dir.create(tsage_path  , recursive = TRUE, showWarnings = FALSE)
dummy = dir.create(tsdbh_path  , recursive = TRUE, showWarnings = FALSE)
dummy = dir.create(tspft_path  , recursive = TRUE, showWarnings = FALSE)
dummy = dir.create(tsmort_path , recursive = TRUE, showWarnings = FALSE)
dummy = dir.create(tsnppo_path , recursive = TRUE, showWarnings = FALSE)
dummy = dir.create(tsauto_path , recursive = TRUE, showWarnings = FALSE)
dummy = dir.create(tsdap_path  , recursive = TRUE, showWarnings = FALSE)
dummy = dir.create(tsdphen_path, recursive = TRUE, showWarnings = FALSE)
dummy = dir.create(tssoil_path , recursive = TRUE, showWarnings = FALSE)
dummy = dir.create(tstheme_path, recursive = TRUE, showWarnings = FALSE)
dummy = dir.create(secy_path   , recursive = TRUE, showWarnings = FALSE)
dummy = dir.create(sens_path   , recursive = TRUE, showWarnings = FALSE)

```





```{r,label="user-pft-fuel"}
user_pftinfo  = TRUE
user_fuelinfo = TRUE

# List of PFTs, in case user_pftinfo is TRUE
n            = 0
pftinfo      = list()
n            = n + 1
pftinfo[[n]] = list( id     = 1
                   , key    = "pft1"
                   , short  = "Oak"
                   , desc   = "Deciduous oak"
                   , colour = "#0072B2"
                   #, parse  ="E*v*g*r*n[Non-h*y*d*r*o]"
                   , parse  = "D*c*i*d[Non-h*y*d*r*o]"
                   )
n            = n + 1
pftinfo[[n]] = list( id     = 2
                   , key    = "pft2"
                   , short  = "Grass"
                   , desc   = "C3 drought decid grass"
                   , colour = "#ffa550"
                   , parse  ="D*e*c*i*d[Non-h*y*d*r*o]"
                   )#end of PFT list


# List of fuels, in case user_fuleinfo is TRUE
n               = 0
fuelinfo        = list()
n               = n + 1
fuelinfo[[n]]   = list( id      = 1
                      , key     = "fuel1"
                      , short   = "Twig"
                      , desc    = "Fuel from twig" 
                      , colour  = "#E69F00"
                      , parse   = "F*u*e*l[T*w*i*g]")
n               = n + 1
fuelinfo[[n]]   = list( id      = 2
                      , key     = "fuel2"
                      , short   = "SBranch"
                      , desc    = "Fuel from small branch" 
                      , colour  = "#56B4E9"
                      , parse   = "F*u*e*l[S*b*r*a*n*c*h]")
n               = n + 1
fuelinfo[[n]]   = list( id      = 3
                      , key     = "fuel3"
                      , short   = "LBranch"
                      , desc    = "Fuel from large branch" 
                      , colour  = "#0072B2"
                      , parse   = "F*u*e*l[L*b*r*a*n*c*h]")
n               = n + 1
fuelinfo[[n]]   = list( id      = 4
                      , key     = "fuel4"
                      , short   = "Trunk"
                      , desc    = "Fuel from trunk" 
                      , colour  = "#D55E00"
                      , parse   = "F*u*e*l[T*r*u*n*k]")
n               = n + 1
fuelinfo[[n]]   = list( id      = 5
                      , key     = "fuel5"
                      , short   = "DeadLeaf"
                      , desc    = "Fuel from dead leaf" 
                      , colour  = "#999999"
                      , parse   = "F*u*e*l[D*l*e*a*f]")
n               = n + 1
fuelinfo[[n]]   = list( id      = 6
                      , key     = "fuel6"
                      , short   = "LiveGrass"
                      , desc    = "Fuel from live grass" 
                      , colour  = "#009E73"
                      , parse   = "F*u*e*l[L*g*r*a*s*s]") #end of fuel list



```






```{r,label="plot-setting"}

gg_device  = c("png")     # Output devices to use (Check ggsave for acceptable formats)
gg_depth   = 600          # Plot resolution (dpi)
gg_ptsz    = 18           # Font size
gg_ptszl   = 26
gg_width   = 17.5         # Plot width (units below)
gg_widthn  = 14.5
gg_height  = 8.5          # Plot height (units below)
gg_units   = "in"         # Units for plot size
gg_screen  = TRUE         # Show plots on screen as well?
gg_tfmt    = "%Y"         # Format for time strings in the time series %Y: 4 DIGITS YEAR %y 2 digits year
gg_ncolours    = 129      # Number of node colours for heat maps.
gg_fleg        = 1./6.    # Fraction of plotting area dedicated for legend
gg_dstat_thick = 0.1      # Thickess for the drought-deciduous status band in the stress overview plot (Value

#ens_colour  =  c("#E69F00", "#56B4E9", "#CC79A7",
                #"#F0E442", "#0000FF", "#00B286")

ens_colour = c("#A0AE6A", "#836B43", "#D68D18", "#437683", "#18B0D6")

top12 <- c("#88CCEE", "#CC6677", "#DDCC77", "#117733", "#332288", "#AA4499", 
           "#44AA99", "#999933", "#882255", "#6699CC", "#888888", "#661100")
cols  = c("#0571B0","#92C5DE","#FFDCD1","#F4A582","#CA0020")

site_colour  = "#000000"

ndevice = length(gg_device)


# Define lower and upper bound for interannual variability ribbon around the mean.
sdev_ribbon  = 1.                  # Standard-deviation equivalent for ribbon 
qlwr_ribbon  = pnorm(-sdev_ribbon) # Lower quantile for ribbon
qupr_ribbon  = pnorm(+sdev_ribbon) # Lower quantile for ribbon
alpha_ribbon = 0.2

```





```{r, label="load-everything"}
source(file.path(util_path,"load.everything.r"),chdir=TRUE)
library(stars)
library(corrplot)

```







```{r,label="time-location-setting"}

nc_conn      = nc_open(case_files[1])

# longitude and latitude
wrf         = nc_open(wrf_path)
lon         = ncvar_get(wrf, "LONGXY")
lat         = ncvar_get(wrf, "LATIXY")
dummy       = nc_close(wrf)
lon_vec     = as.vector(lon)
lat_vec     = as.vector(lat)
nlon        = length(lon_vec)
nlat        = length(lat_vec)

# time
time_val     = ncvar_get(nc_conn,"time")
time_use     = time_val
nstep        = length(time_use)
model_yra    = 1951
n_yr         = round((tail(time_use,1)-head(time_use,1))/365,digits=0)
yr_start     = model_yra + floor(time_use[1]/365)
yr_end       = yr_start + n_yr-1
simu_time    = c(yr_start:yr_end)
nc_year      = as.numeric(yr_start:yr_end)
nc_month     = as.numeric(1:12)
yr_ens       = rep(nc_year,each=nlon*12)
month_yr     = rep(nc_month,each=nlon)
month_ens    = rep(month_yr,times=n_yr)
ens_tstamp   = make_datetime(year=yr_ens,month=month_ens,day=1L)
ntstamp_ens = length(ens_tstamp)
dummy        = nc_close(nc_conn)

```








```{r,label='redefine-read-group',message=FALSE,results='hide'}
read_dbh   = FALSE
read_age   = FALSE
read_fuel  = TRUE
read_hlm2d = FALSE
read_phen  = FALSE
read_soi   = FALSE

fates_var  = c( "fates_basalarea","fates_crownarea",	"fates_ddbh_canopy",
               "fates_ddbh_ustory","fates_fuel_amount",
               "fates_gpp","fates_lai_canopy","fates_lai_ustory",
               "fates_mortality_background","fates_mortality_cstarv",
               "fates_mortality_freezing","fates_mortality_fire",
               "fates_mortality_hydraulic","fates_mortality_termination",
               "fates_nplant_canopy","fates_nplant_ustory",
               "fates_patcharea","fates_vegc_aboveground","fates_lai","fates_nplant") 

fatesvar_new  = fatesvar                   %>% 
                filter(vnam %in% fates_var)
nfatesvar_new = length(fatesvar_new$vnam)

### define variable to read for size class only
szpf_var = c("fates_basalarea","fates_nplant","fates_vegc_aboveground","fates_ddbh_canopy",
             "fates_ddbh_ustory","fates_nplant_canopy","fates_nplant_ustory")
fatesvar_szpf = fatesvar     %>%  filter(vnam %in% szpf_var)
nszpf_read    = length(fatesvar_szpf$vnam)
               

```






```{r,label="dimension-setting"}
if ("nc_conn" %in% ls()){dummy = nc_close(nc_conn); rm(nc_conn)}

nc_dlist <- list()
nc_vlist <- list()

nc_conn  = nc_open(filename=case_files[1])
nc_nvars = nc_conn$nvars
nc_ndims = nc_conn$ndims
nc_dlist = rep(NA_character_,times=nc_ndims)
nc_vlist = rep(NA_character_,times=nc_nvars)
for (d in sequence(nc_ndims)) nc_dlist[d] = nc_conn$dim[[d]]$name
for (v in sequence(nc_nvars)) nc_vlist[v] = nc_conn$var[[v]]$name

#---~---
# Gather dimension information, then initialize matrices
#---~---

if(read_age){
  # List of age classes
idxage   = match("fates_levage",nc_dlist)
if (is.finite(idxage)){
   ages     = nc_conn$dim[[idxage]]$vals
   nages    = nc_conn$dim[[idxage]]$len

   ageinfo  = tibble( id      = sequence(nages)
                    , age_lwr = ages
                    , age_upr = c(ages[-1],Inf)
                    , key     = sprintf("age_%3.3i",ages)
                    , desc    = c( paste0("paste(paste(",age_lwr[-nages],"<= A*g*e)<",age_upr[-nages],"*y*r)")
                                 , paste0("paste( A*g*e >=",age_upr[nages],"*y*r)")
                                 )#end c
                    , labs    = c( paste0("paste(",age_lwr[-nages],"-",age_upr[-nages],")")
                                 , paste0("paste(",age_lwr[ nages],"-infinity)")
                                 )#end c
                    , colour  = viridis(nages,option="D",direction=-1)
                    )#end tibble
}else{
   ageinfo  = tibble( id      = integer(0L)
                    , age_lwr = numeric(0L)
                    , age_upr = numeric(0L)
                    , key     = character(0L)
                    , desc    = character(0L)
                    , labs    = character(0L)
                    , colour  = character(0L)
                    )#end tibble
}#end if (is.na(idxage))

# Set number of age classes
nages = nrow(ageinfo)

#---~---
#retrieve age-class variables
#---~---
nc_byage = nc_vlist[grepl(pattern="_AP$",x=nc_vlist)]
nc_pref  = tolower(gsub(pattern="_AP$",replacement="",x=nc_byage))
nc_keep  = nc_pref %in% fatesvar_new$vnam & (! duplicated(nc_pref))
no_byage = nc_byage[! nc_keep]
nc_byage = nc_byage[  nc_keep]
nbyage   = length(nc_byage)

}



# List of size classes
if(read_dbh){
  idxdbh   = match("fates_levscls",nc_dlist)
if (is.finite(idxdbh)){
   dbhs     = nc_conn$dim[[idxdbh]]$vals
   ndbhs    = nc_conn$dim[[idxdbh]]$len
   dbhinfo  = tibble( id      = sequence(ndbhs)
                    , dbh_lwr = dbhs
                    , dbh_upr = c(dbh_lwr[-1],dbh_lwr[ndbhs]+2*max(diff(dbh_lwr)))
                    , dbh     = 0.5 * (dbh_lwr + dbh_upr)
                    , key     = sprintf("dbh_%3.3g",dbh_lwr)
                    , desc    = c( paste0("paste(paste(",dbh_lwr[-ndbhs],"<=D*B*H)<",dbh_upr[-ndbhs],"*c*m)")
                                 , paste0("paste( D*B*H >=",dbh_lwr[ndbhs],"*c*m)")
                                 )#end c
                    , labs    = c( paste0("paste(",dbh_lwr[-ndbhs],"-",dbh_upr[-ndbhs],")")
                                 , paste0("paste(",dbh_lwr[ ndbhs],"-infinity)")
                                 )#end dbhlabs
                    , colour   = magma(ndbhs,direction=1)
                    )#end tibble
}else{
   dbhinfo  = tibble( id      = integer(0L)
                    , dbh_lwr = numeric(0L)
                    , dbh_upr = numeric(0L)
                    , key     = character(0L)
                    , desc    = character(0L)
                    , labs    = character(0L)
                    , colour  = character(0L)
                    )#end tibble
}#end if (is.finite(idxdbh))

# Set number of size classes
ndbhs = nrow(dbhinfo)


# by size class and whether total LAI can be calculated given canopy
# and under-story LAI

is_size   = grepl(pattern="_SZ$",x=nc_vlist) | grepl(pattern="_SZPF$",x=nc_vlist)
nc_bydbh  = nc_vlist[is_size]
nc_pref   = gsub(pattern="_SZ$",replacement="",x=nc_bydbh)
nc_pref   = gsub(pattern="_SZPF$",replacement="",x=nc_pref )
nc_pref   = tolower(nc_pref)
nc_keep   = (nc_pref %in% fatesvar_szpf$vnam) & (! duplicated(nc_pref))
no_bydbh  = nc_bydbh[! nc_keep]
nc_bydbh  = unique(nc_bydbh[nc_keep])

vardbh_last = rep(x=FALSE,times=nszpf_read)
for (v in which(fatesvar_szpf$is_upc)){
   nc_vnow      = toupper(fatesvar_szpf$vnam[v])
   nc_vnow_scls = paste0(nc_vnow,"_SZ")
   nc_vund_scls = paste0(nc_vnow,"_USTORY_SZ")
   nc_vcan_scls = paste0(nc_vnow,"_CANOPY_SZ")
   nc_vnow_scpf = paste0(nc_vnow,"_SZPF")
   nc_vund_scpf = paste0(nc_vnow,"_USTORY_SZPF")
   nc_vcan_scpf = paste0(nc_vnow,"_CANOPY_SZPF")

   # Check whether this variable can be derived from understorey+canopy (and needs to be).
   if ( all(c(nc_vund_scls,nc_vcan_scls) %in% nc_bydbh ) && (! nc_vnow_scls %in% nc_bydbh) ){
      nc_bydbh       = unique(c(nc_bydbh,nc_vnow_scls))
      vardbh_last[v] = TRUE
   }else if ( all(c(nc_vund_scpf,nc_vcan_scpf) %in% nc_bydbh ) && (! nc_vnow_scpf %in% nc_bydbh) ){
      nc_bydbh       = unique(c(nc_bydbh,nc_vnow_scpf))
      vardbh_last[v] = TRUE
   }#end if ( all(c(nc_vund_scls,nc_vcan_scls) %in% nc_bydbh ) && (! nc_vnow_scls %in% nc_bydbh) )
}#end for (v in which(fatesvar$is_upc))


nbydbh    = length(nc_bydbh)
n_vardbh_last = sum(vardbh_last)

}


# List of PFT classes (only if not using user-defined classes).
idxpft   = match("fates_levpft",nc_dlist)
if (! is.finite(idxpft)){
   # PFT index not found. Skip PFTs altogether.
   pftinfo = tibble( id               = numeric(0L)
                   , key              = character(0L)
                   , short            = character(0L)
                   , desc             = character(0L)
                   , colour           = character(0L)
                   , stringsAsFactors = FALSE
                   )#end data.table
}else if (! user_pftinfo){
   # Select all PFTs available
   pftids  = nc_conn$dim[[idxpft]]$vals
   npftids = nc_conn$dim[[idxpft]]$len

   # Build tibble with all the PFTs.
   pftinfo = tibble( id               = pftids
                   , key              = sprintf("pft%2.2i" ,pftids)
                   , short            = sprintf("PFT%2.2i" ,pftids)
                   , desc             = sprintf("PFT %2.2i",pftids)
                   , colour           = brewer.pal(n=npftids,name="PuBuGn")
                   , stringsAsFactors = FALSE
                   )#end tibble
}else if (! is_tibble(pftinfo)){
   # Convert user-defined pftinfo to a "tibble" object
   pftinfo  = do.call(what=rbind,args=lapply(X=pftinfo,FUN=as_tibble,stringsAsFactors=FALSE))
}#end if (! is.finite(idxpft))

# Set number of PFTs (active PFTs only) 
npfts = nrow(pftinfo)

is_pft    = grepl(pattern="_SZPF$",x=nc_vlist) | grepl(pattern="_PF$",x=nc_vlist) | grepl(pattern="_$",x=nc_vlist) #for truncated hydro_cflux
nc_bypft  = nc_vlist[is_pft]
nc_pref   = gsub(pattern="_SZPF$",replacement="",x=nc_bypft )
nc_pref   = gsub(pattern="_PF$",replacement="", x=nc_pref  )
nc_pref   = gsub(pattern="_$",replacement="",x=nc_pref     )
nc_pref   = tolower(nc_pref)
nc_keep   = nc_pref %in% fatesvar_new$vnam & (! duplicated(nc_pref))
no_bypft  = nc_bypft[! nc_keep]
nc_bypft  = unique(nc_bypft[nc_keep])

varpft_last = rep(x=FALSE,times=nfatesvar_new)
for (v in which(fatesvar_new$is_upc)){
   nc_vnow      = toupper(fatesvar_new$vnam[v])
   nc_vnow_scpf = paste0(nc_vnow,"_SZPF")
   nc_vund_scpf = paste0(nc_vnow,"_USTORY_SZPF")
   nc_vcan_scpf = paste0(nc_vnow,"_CANOPY_SZPF")

   # Check whether this variable can be derived from understorey+canopy (and needs to be).
   if ( all(c(nc_vund_scpf,nc_vcan_scpf) %in% nc_bypft)  && (! nc_vnow_scpf %in% nc_bypft) ){
      nc_bypft       = unique(c(nc_bypft,nc_vnow_scpf))
      varpft_last[v] = TRUE
   }#end if ( all(c(nc_vund_scls,nc_vcan_scls) %in% nc_bydbh ) && (! nc_vnow_scls %in% nc_bydbh) )
}#end for (v in which(fatesvar$is_upc))

nbypft    = length(nc_bypft)
n_varpft_last = sum(varpft_last)


# List of fuel classes
if(read_fuel){
  idxfuel   = match("fates_levfuel",nc_dlist)
  if (! is.finite(idxfuel)){
   # fuel index not found. Skip Fuels altogether.
     fuelinfo = tibble( id               = numeric(0L)
                      , key              = character(0L)
                      , short            = character(0L)
                      , desc             = character(0L)
                      , colour           = character(0L)
                      , stringsAsFactors = FALSE
                      )#end data.table
}else if (! user_fuelinfo){
   # Select all fuels available
   fuelids  = nc_conn$dim[[idxfuel]]$vals
   nfuelids = nc_conn$dim[[idxfuel]]$len

   # Build tibble with all the fuels.
   fuelinfo = tibble( id               = fuelids
                    , key              = sprintf("Fuel %2.2i" ,fuelids)
                    , short            = sprintf("Fuel %2.2i" ,fuelids)
                    , desc             = sprintf("Fuel %2.2i" ,fuelids)
                    , colour           = brewer.pal(n=nfuelids,name="PuBuGn")
                    , stringsAsFactors = FALSE
                   )#end tibble
}else if (! is_tibble(fuelinfo)){
   # Convert user-defined fuelinfo to a "tibble" object
   fuelinfo  = do.call(what=rbind,args=lapply(X=fuelinfo,FUN=as_tibble,stringsAsFactors=FALSE))
}#end if (! is.finite(idxfuel))
nfuels = nrow(fuelinfo)
  
}

# by fuel class
nc_byfuel = nc_vlist[grepl(pattern="_FC$",x=nc_vlist)]
nc_pref   = tolower(gsub(pattern="_FC$",replacement="",x=nc_byfuel))
nc_keep   = nc_pref %in% fatesvar_new$vnam & (! duplicated(nc_pref))
no_byfuel = nc_byfuel[! nc_keep]
nc_byfuel = nc_byfuel[  nc_keep]
nbyfuel   = length(nc_byfuel)


# List of soil layer
if(read_soi){
  idxsoi   = match("levgrnd",nc_dlist)
if (is.finite(idxsoi)){
   sois     = nc_conn$dim[[idxsoi]]$vals
   nsois    = nc_conn$dim[[idxsoi]]$len
   soikeys  = sprintf("sl_%3.3f",sois)
   soilabs  = c( paste0("paste(",sois[-nsois],"-",sois[-1],")")
               , paste0("paste(",sois[nsois],"-infinity)")
               )#end dbhlabs
   soilabs  = parse(text=soilabs)
}else{
   sois    = numeric(0L)
   nsois   = 0L
   soikeys = character(0L)
   soilabs = character(0L)
}#end if (is.finite(idxdbh))
}


#---~---
#    Retrieve all "drought deciduous phenology variables
#---~---

if(read_phen){
  nc_pref   = tolower(x=nc_vlist)
  nc_keep   = nc_pref %in% dphenvar$vorig
  no_dphen  = nc_vlist[! nc_keep]
  nc_dphen  = nc_vlist[  nc_keep]

# Tally the total number of drought phenology variables.
  ndphen = length(nc_dphen)
  
}



# all 1D and 2D variables variables
nc_pref   = tolower(x=nc_vlist)
nc1d_keep = nc_pref %in% hlm1dvar$vnam
no_hlm1d  = nc_vlist[! nc1d_keep]
nc_hlm1d  = nc_vlist[  nc1d_keep]
nc2d_keep = nc_pref %in% hlm2dsoi$vnam
no_hlm2d  = nc_vlist[! nc2d_keep]
nc_hlm2d  = nc_vlist[  nc2d_keep]

# Check whether to append "evapotranspiration"
if (  ( all(c("QSOIL","QVEGT","QVEGE") %in% nc_hlm1d) ) && (! "QEVTR" %in% nc_hlm1d) ){
   nc_hlm1d = unique(c(nc_hlm1d,"QEVTR"))
   etr_last = TRUE
}else{
   etr_last = FALSE
}#end if (  ( all(c("QSOIL","QVEGT","QVEGE") %in% nc_hlm1d) ) && (! "QEVTR" %in% nc_hlm1d) )

# Check whether to append ecosystem respiration
if (  ( all(c("FATES_AUTORESP","FATES_HET_RESP") %in% nc_hlm1d) ) && (! "ER" %in% nc_hlm1d) ){
   nc_hlm1d = unique(c(nc_hlm1d,"ER"))
   er_last  = TRUE
}else{
   er_last  = FALSE
}#end if (  ( all(c("AR","HR") %in% nc_hlm1d) ) && (! "ER" %in% nc_hlm1d) )

# Find number of host land model variables
nhlm1d    = length(nc_hlm1d)
nhlm2d    = length(nc_hlm2d)

#---~---
#Initialize variable matrix
#---~---

#by age class
#byage = list()
#for(c in sequence(ncase)){
#  case     = as.character(case_names[c])
  
#  for(a in sequence(nbyage)){
#  nc_nvnow                   = nc_byage[a]
#  nc_pref                    = tolower(gsub(pattern="_AP$",replacement="",x=nc_nvnow))
#  f                          = match(nc_pref,fatesvar$vnam)
#  f_vnam                     = fatesvar$vnam[f]
#  byage[[case]][[f_vnam]]    = matrix(data=NA_real_,nrow=ntstamp_ens,ncol=nages,dimnames=list(NULL,ageinfo$key)) 
#  } #end of a loop
#} #end of c(in sequence(ncase))


# by size class
if(read_dbh){
  bydbh = list()
  for(c in sequence(ncase)){
  case     = as.character(case_names[c])
  for(d in sequence(nbydbh)){
  nc_nvnow        = nc_bydbh[d]
  nc_pref         = gsub(pattern="_SZ$",replacement="",x=nc_nvnow)
  nc_pref         = gsub(pattern="_SZPF$",replacement="",x=nc_pref )
  nc_pref         = tolower(nc_pref)
  f               = match(nc_pref,fatesvar_szpf$vnam)
  f_vnam          = fatesvar_szpf$vnam[f]
  bydbh[[case]][[f_vnam]] = matrix(data=NA_real_,nrow=ntstamp_ens,ncol=ndbhs,dimnames=list(NULL,dbhinfo$key))
   } # end of d
}
  
}



# by pft

bypft = list()
for(c in sequence(ncase)){
  case     = as.character(case_names[c])
  for(p in sequence(nbypft)){
  nc_nvnow        = nc_bypft[p]
  nc_pref         = gsub(pattern="_PF$",replacement="",x=nc_nvnow  )
  nc_pref         = gsub(pattern="_SZPF$",replacement="",x=nc_pref )
  nc_pref         = gsub(pattern="_$",replacement="",x=nc_pref )
  nc_pref         = tolower(nc_pref)
  f               = match(nc_pref,fatesvar_new$vnam)
  f_vnam          = fatesvar_new$vnam[f]
  bypft[[case]][[f_vnam]] = matrix(data=NA_real_,nrow=ntstamp_ens,ncol=npfts,
                           dimnames=list(NULL,pftinfo$key)) 
  } # end of p
}

# Initialize list of variables by fuel class

byfuel = list()
for(c in sequence(ncase)){
  case     = as.character(case_names[c])
  
  for(f in sequence(nbyfuel)){
  nc_nvnow                   = nc_byfuel[f]
  nc_pref                    = tolower(gsub(pattern="_FC$",replacement="",x=nc_nvnow))
  f                          = match(nc_pref,fatesvar_new$vnam)
  f_vnam                     = fatesvar_new$vnam[f]
  byfuel[[case]][[f_vnam]] = matrix(data=NA_real_,nrow=ntstamp_ens,ncol=nfuels,dimnames=list(NULL,fuelinfo$key)) 
  } #end of a loop
} #end of c(in sequence(ncase))


# Initialise list of variables by soil layer
#dphen = list()
#for(c in sequence(ncase)){
#  case     = as.character(case_names[c])
#  for (p in sequence(ndphen)){
#   nc_nvnow        = nc_dphen[p]
#   nc_pref         = tolower(nc_nvnow)
#   f               = match(nc_pref,dphenvar$vorig)
#   f_vnam          = dphenvar$vnam[f]
#   dphen[[case]][[f_vnam]] = matrix(data=NA_real_,nrow=ntstamp_ens,
#                                    ncol=npfts,dimnames=list(NULL,pftinfo$key))
#  }#end for (p in sequence(ndphen))
#}



# Initialize 1D variables available at the HLM
hlm1d = as_tibble( matrix( data     = NA_real_
                         , nrow     = ntstamp_ens*ncase
                         , ncol     = nhlm1d+1
                         , dimnames = list(NULL,tolower(c(nc_hlm1d,"CASE")))
                         )
                 )

# Initialize soil water status related variables available at the HLM for each soil layer into hlm2dvar
#hlm2d = list()
#for(c in sequence(ncase)){
#  case     = as.character(case_names[c])
  
#  for(m in sequence(nhlm2dsoi)){
#  nc_nvnow        = nc_hlm2d[m]
#  nc_pref         = tolower(nc_nvnow)
#  h               = match(nc_pref,hlm2dsoi$vnam)
#  h_vnam          = hlm2dsoi$vnam[h]
#  hlm2d[[case]][[h_vnam]] = matrix(data=NA_real_,nrow=ntstamp_ens*ncase,ncol=nsois,dimnames=list(NULL,soikeys))
#  }
#}


# Load soil layers
# do a loop to avoid hard-coding here 
#cat0("   - Load soil information")
#slayer = tibble()

#for (c in sequence(ncase)) {
#case     = as.character(case_names[c])
#slayer_now = tibble( zsoi   = c(unlist(ncvar_get(nc=nc_conn, varid='ZSOI',collapse_degen=TRUE))) 
#                   , dzsoi  = c(unlist(ncvar_get(nc=nc_conn,varid='DZSOI',collapse_degen=TRUE ))) 
#                   , bsw    = c(unlist(ncvar_get(nc=nc_conn,varid='BSW',collapse_degen=TRUE )))  
#                   , hksat  = c(unlist(ncvar_get(nc=nc_conn,varid='HKSAT',collapse_degen=TRUE ))) 
#                   , sucsat = c(unlist(ncvar_get(nc=nc_conn,varid='SUCSAT',collapse_degen=TRUE )))  
#                   , watsat = c(unlist(ncvar_get(nc=nc_conn,varid='WATSAT',collapse_degen=TRUE )))  
#                   , case   = c(rep_along(zsoi, case))
#               )#end tibble
#   slayer = rbind(slayer, slayer_now)
#   rm(slayer_now)
#}


# Load indices
scls   = ncvar_get(nc=nc_conn,varid='fates_scmap_levscpf')
pft    = ncvar_get(nc=nc_conn,varid='fates_pftmap_levscpf')
fuel   = ncvar_get(nc=nc_conn,varid='fates_levfuel')
index_scpf = tibble( scls    
                   , pft    
                   )#end data.table;  
rm(scls,pft)
# Close connection
dummy   = nc_close(nc_conn) 

```






```{r,label="retrieve-values",message=FALSE,results='hide'}
if ("nc_conn" %in% ls()){dummy = nc_close(nc_conn); rm(nc_conn)}

#for(n in sequence(ntstamp_each)){

 # w_month      = month(each_tstamp[w])  
 # w_year       = year(each_tstamp[w])
  
  # Find conversion factors for monthly variables.
  
   cmon.day = days_in_month(ens_tstamp)
   cmon.hr  = day.hr  * cmon.day
   cmon.min = day.min * cmon.day
   cmon.sec = day.sec * cmon.day
  

# Open NetCDF connection and retrieve variable names
for(c in sequence(ncase)){
  case     = case_names[c]
  nc_conn  = nc_open(case_files[c])
  nc_nvars = nc_conn$nvars
  nc_vlist = rep(NA_character_,times=nc_nvars)
  for (v in sequence(nc_nvars)) nc_vlist[v] = nc_conn$var[[v]]$name


# Read variables by age, and assign current values to the matrix.  
#  for (a in sequence(nbyage)){
#    nc_nvnow            = nc_byage[a]
#    nc_pref             = tolower(gsub(pattern="_AP$",replacement="",x=nc_nvnow))
#    f                   = match(nc_pref,fatesvar$vnam)
#    f_vnam              = fatesvar$vnam[f]
#    f_add0              = eval(parse(text=fatesvar$add0[f]))
#    f_mult              = eval(parse(text=fatesvar$mult[f]))
#    nc_dat              = ncvar_get(nc=nc_conn,varid=nc_nvnow,collapse_degen = TRUE)
   #subset time to only include results from 2000 to 2020
#    nc_dat              = nc_dat[,,,229:480] 
   #reform the 3d array so when convert to matrix, rows are time by ensemble members, columns are age class 
#    ncm_dat             = aperm(nc_dat, c(2,3,4,1)) 
#    dim(ncm_dat)        = c(ntstamp_ens,nages)
#    nc_dat              = f_add0 + f_mult * ncm_dat
#    byage[[case]][[f_vnam]][,]   = nc_dat
#}#end for (a in sequence(nbyage))  

#---~--- 
   #   Read variables by size, and assign current values to the matrix.  In case LAI
   # is in the list and it is the last variable, we calculate it from canopy and under
   # story, after loading all variables
   #---~---
if(read_dbh){
  for (d in sequence(nbydbh-n_vardbh_last)){
      nc_nvnow            = nc_bydbh[d]
      nc_pref             = gsub(pattern="_SZPF$",replacement="",x=nc_nvnow)
      nc_pref             = gsub(pattern="_SZ$",replacement="",x=nc_pref )
      nc_pref             = tolower(nc_pref)
      f                   = match(nc_pref,fatesvar_szpf$vnam)
      f_vnam              = fatesvar_szpf$vnam[f]
      f_add0              = eval(parse(text=fatesvar_szpf$add0[f]))
      f_mult              = eval(parse(text=fatesvar_szpf$mult[f]))
      f_aggr              = match.fun(fatesvar_szpf$aggr[f])
      nc_dat              = ncvar_get(nc=nc_conn,varid=nc_nvnow)
      nc_dat              = nc_dat[,,,] 
      ncm_dat             = aperm(nc_dat,c(1,2,4,3))
      szpf_dim            = dim(ncm_dat)[4]
      dim(ncm_dat)        = c(ntstamp_ens,szpf_dim)
      nc_dat              = f_add0 + f_mult * ncm_dat
      bydbh[[case]][[f_vnam]][,]  = nc_dat[,1:14]    # we only grab size-class variable for oak [PFT 1]
   }#end for (d in sequence(nbydbh-laidbh_last))

#    Loop through variables to be added last
   for (v in which(vardbh_last)){
      # Retrieve variable and build understory and canopy variables.
      v_vnam = fatesvar_szpf$vnam[v]
      v_vund = paste0(v_vnam,"_ustory")
      v_vcan = paste0(v_vnam,"_canopy")

      # Aggregate data
      bydbh[[case]][[v_vnam]] = bydbh[[case]][[v_vund]] + bydbh[[case]][[v_vcan]]
  }#end for (v in which(vardbh_last))
  
}

  
 

#---~---
# read variables by PFT
#---~---
  for (p in sequence(nbypft-n_varpft_last)){
      # Load variable information
      nc_nvnow = nc_bypft[p]
      is_szpf  = grepl(pattern="_SZPF$",x=nc_nvnow)
      nc_pref  = gsub(pattern="_SZPF$",replacement="",x=nc_nvnow)
      nc_pref  = gsub(pattern="_PF$",replacement="",x=nc_pref )
      nc_pref  = gsub(pattern="_$",replacement="",x=nc_pref )
      nc_pref  = tolower(nc_pref)
      f        = match(nc_pref,fatesvar_new$vnam)
      f_vnam   = fatesvar_new$vnam[f]
      f_add0   = eval(parse(text=fatesvar_new$add0[f]))
      f_mult   = eval(parse(text=fatesvar_new$mult[f]))
      f_aggr   = match.fun(fatesvar_new$aggr[f])
      nc_dat   = ncvar_get(nc=nc_conn,varid=nc_nvnow,collapse_degen = FALSE)
      trd_dim  = dim(nc_dat)[3] # this is the pft * size dim
      sum_dim  = dim(nc_dat)[1]*dim(nc_dat)[2]*dim(nc_dat)[4]
      if(is_szpf){
      ncm_dat        = aperm(nc_dat,c(1,2,4,3))
      dim(ncm_dat)   = c(sum_dim,trd_dim)
      ncm_dat        = f_add0 + f_mult * ncm_dat
      #nc_aggr        = rowSums(ncm_dat)
      p1             = rowSums(ncm_dat[,2:14]) # for oak we ignore the first size class
      p2             = rowSums(ncm_dat[,15:28])
      nc_aggr        = rbind(p1,p2)
      nc_aggr        = aperm(nc_aggr,c(2,1))
      bypft[[case]][[f_vnam]][] = nc_aggr
      }else{
      nc_dat      = aperm(nc_dat,c(1,2,4,3))
      dim(nc_dat) = c(ntstamp_ens,trd_dim)
      nc_dat = f_add0 + f_mult * nc_dat
      bypft[[case]][[f_vnam]][] = nc_dat
      }
}#end for (p in sequence(nbypft-n_varpft_last))

# Loop through variables to be added last
  for (v in which(varpft_last)){
      # Retrieve variable and build understory and canopy variables.
    v_vnam = fatesvar_new$vnam[v]
    v_vund = paste0(v_vnam,"_ustory")
    v_vcan = paste0(v_vnam,"_canopy")

  # Aggregate data
   bypft[[case]][[v_vnam]] = bypft[[case]][[v_vund]] + bypft[[case]][[v_vcan]]
   }#end for (v in which(varpft_last))
  

   # Read variables by fuel class, and assign current values to the matrix.  
# for (f in sequence(nbyfuel)){
#    nc_nvnow            = nc_byfuel[f]
#    nc_pref             = tolower(gsub(pattern="_FC$",replacement="",x=nc_nvnow))
#    f                   = match(nc_pref,fatesvar$vnam)
#    f_vnam              = fatesvar$vnam[f]
#    f_add0              = eval(parse(text=fatesvar$add0[f]))
#    f_mult              = eval(parse(text=fatesvar$mult[f]))
#    nc_dat              = ncvar_get(nc=nc_conn,varid=nc_nvnow,collapse_degen = TRUE)
   #subset time to only include results from 2000 to 2020
#    nc_dat              = nc_dat[,,,229:480] 
   #reform the 3d array so when convert to matrix, rows are time by ensemble members, columns are fuel class 
#    ncm_dat             = aperm(nc_dat, c(1,2,4,3)) 
#    dim(ncm_dat)        = c(ntstamp_ens, nfuels)
#    ncm_dat              = f_add0 + f_mult * ncm_dat
#    byfuel[[case]][[f_vnam]][,]   = ncm_dat
#}#end for (f in sequence(nbyfuel))  


 # Read drought-deciduous phenology variables (by PFT), and assign current values to the matrix.
   # We use "rev" because the first soil layer is the deepest for the R output.
#   for (p in sequence(ndphen)){
#      nc_nvnow            = nc_dphen[p]
#      nc_pref             = tolower(nc_nvnow)
#      f                   = match(nc_pref,dphenvar$vorig)
#      f_vnam              = dphenvar$vnam[f]
#      f_add0              = eval(parse(text=dphenvar$add0[f]))
#      f_mult              = eval(parse(text=dphenvar$mult[f]))
#      nc_dat              = ncvar_get(nc=nc_conn,varid=nc_nvnow,collapse_degen = TRUE)
#      dphen[[case]][[f_vnam]][] = f_add0 + f_mult * rev(nc_dat)
      #dphen[[case]][[f_vnam]][,] = f_add0 + f_mult * rev(nc_dat)
#   }#end for (a in sequence(nbyage))


# Read 1D variables
 
  for (v in sequence(nhlm1d-etr_last-er_last)){
  
  nc_nvnow            = nc_hlm1d[v]
  nc_pref             = tolower(x=nc_nvnow)
  h                   = match(nc_pref,hlm1dvar$vnam)
  h_vnam              = hlm1dvar$vnam[h]
  h_add0              = eval(parse(text=hlm1dvar$add0[h]))
  h_mult              = eval(parse(text=hlm1dvar$mult[h]))
  nc_dat              = ncvar_get(nc=nc_conn,varid=nc_nvnow,collapse_degen = TRUE)
  ncm_dat             = matrix(nc_dat, ncol=1)
  
  if (c==1){
         hlm1d[[h_vnam]][1:ntstamp_ens]   = h_add0 + h_mult * ncm_dat
         hlm1d$case[1:ntstamp_ens]        = case
     }else{
         start = (c-1)*ntstamp_ens+1
         end   = c*ntstamp_ens
         hlm1d[[h_vnam]][start:end] = h_add0 + h_mult * ncm_dat
         hlm1d$case[start:end]      = case
         }
  
}#for (h in sequence(nhlm1d-etr_last-et_last))
  
 # Find total ET.
   if (etr_last){
      if(c==1){
         hlm1d$qevtr[1:ntstamp_ens] =  hlm1d$qvege[1:ntstamp_ens] + 
                                       hlm1d$qvegt[1:ntstamp_ens] + 
                                       hlm1d$qsoil[1:ntstamp_ens]
      }else{
         start = (c-1)*ntstamp_ens+1
         end   = c*ntstamp_ens
         hlm1d$qevtr[start:end] = hlm1d$qvege[start:end] + 
                                  hlm1d$qvegt[start:end] + 
                                  hlm1d$qsoil[start:end]
      }
   }#end if (etr_last)

   # Find total ET.
   if (er_last){
      if(c==1){
         hlm1d$er[1:ntstamp_ens] =  hlm1d$fates_autoresp[1:ntstamp_ens] + 
                                    hlm1d$fates_het_resp[1:ntstamp_ens]
      }else{
         start = (c-1)*ntstamp_ens+1
         end   = c*ntstamp_ens
         hlm1d$er[start:end] = hlm1d$fates_autoresp[start:end] + 
                               hlm1d$fates_het_resp[start:end] 
               }
   }#end if (er_last)


# Read 2D variables
#for (m in sequence(nhlm2d)){ 
#  nc_nvnow            = nc_hlm2d[m]
#  nc_pref             = tolower(x=nc_nvnow)
#  h                   = match(nc_pref,hlm2dsoi$vnam)
#  h_vnam              = hlm2dsoi$vnam[h]
#  h_add0              = eval(parse(text=hlm2dsoi$add0[h]))
#  h_mult              = eval(parse(text=hlm2dsoi$mult[h]))
#  nc_dat              = ncvar_get(nc=nc_conn,varid=nc_nvnow,collapse_degen = TRUE)
#  nc_dat              = nc_dat[,181:432,] 
#  ncm_dat             = aperm(nc_dat, c(2,3,1))
#  dim(ncm_dat)        = c(ntstamp_each*ninst, nsois)
#  nc_dat              = h_add0 + h_mult * ncm_dat
#  hlm2d[[case]][[h_vnam]][,]  = nc_dat
#}#end of for (m in sequence(nhlm2d))
# Close connection
dummy   = nc_close(nc_conn)

  
} #end of for (c in sequence(ncase))

## add time for hlm1d
hlm1d$time = rep(ens_tstamp,times=ncase)
hlm1d$lon  = rep(lon_vec, times=nstep*ncase)
hlm1d$lat  = rep(lat_vec, times=nstep*ncase)
fwrite(hlm1d,file.path(main_path,paste0(file_desc, "_hlm1d.csv")))

```






```{r,label="scale-age",message=FALSE,results='hide'}
cat0(" + Rescale variables by age class.")

for(c in sequence(ncase)){
  case = case_names[c]
  for (a in sequence(nbyage)){
#--- Match variables
  f        = match(names(byage[[case]])[a],fatesvar$vnam)
  f_vnam   = fatesvar$vnam [f]
  f_desc   = fatesvar$desc [f]
  f_stack  = fatesvar$stack[f]
   #---~---

   #--- Proceed only if the variable is stacked and not the patch area.
  if (f_stack && (! (f_vnam %in% "fates_patcharea"))){
      cat0("   - ",f_desc,".")
      byage[[case]][[f_vnam]] = byage[[case]][[f_vnam]] * byage[[case]]$fates_patcharea
  }#end if (f_stack && (! (f_vnam %in% "patch_area")))
}#end for (a in sequence(nbyage))
  
}



```







```{r,label="melt-data",message=FALSE,results='hide'}
#cat0(" + Turn age-dependent matrices into tibble objects.")
#age_melt  = NULL

#for (c in sequence(ncase)){
#  case = case_names[c]
#  now_agemelt = NULL
#  for (a in sequence(nbyage)){
   #--- Match variables.
#  f        = match(names(byage[[case]])[a],fatesvar$vnam)
#  f_vnam   = fatesvar$vnam[f]
#  f_desc   = fatesvar$desc[f]
#  cat0("   - ",f_desc,".")
      
#--- Create molten data table for this variable.  
#  now_age           = as_tibble(byage[[case]][[f_vnam]])
#  now_age$time      = rep(each_tstamp, time=ninst)
#  now_age$case      = rep_along(time, x = case)
#  now_age$ens_label = rep(ens_labels, each = ntstamp_each)
#  now_melt          = as_tibble(reshape2::melt(data=now_age,id.vars=c("case","ens_label",       
#                              "time"),variable.name="age",value.name=f_vnam))
      #--- Merge data table
  
#  if (is.null(now_agemelt)){
#    now_agemelt = now_melt
#  }else{
#        now_agemelt = as_tibble(merge(x=now_agemelt,y=now_melt,by=c("case","ens_label","time","age"),all=TRUE))
#  }#end if (is.null(age_melt))
  
#  }#end for (a in sequence(nbyage))
#age_melt = rbind(age_melt,now_agemelt)
#}



#byage = age_melt 
#rm(age_melt)
#invisible(gc())

#cat0(" + Turn size-dependent matrices into data tables.")
dbh_melt = NULL
for (c in sequence(ncase)){
  case = case_names[c]
for (d in sequence(nbydbh)){
  f        = match(names(bydbh[[case]])[d],fatesvar_szpf$vnam)
  f_vnam   = fatesvar_szpf$vnam[f]
  f_desc   = fatesvar_szpf$desc[f]
  cat0("   - ",f_desc,".")
   
  now_dbh           = as_tibble(bydbh[[case]][[d]])
  now_dbh$lon       = rep(lon_vec,times=nstep)
  now_dbh$lat       = rep(lat_vec,times=nstep)
  now_dbh$time      = ens_tstamp
  now_dbh$case      = rep_along(time, x = case)
  now_melt          = as_tibble(reshape2::melt(data=now_dbh,id.vars=c("case", "lon","lat","time"),variable.name="dbh",value.name=f_vnam))
  
  if (is.null(dbh_melt)){
      dbh_melt = now_melt
      }else{
      #dbh_melt = merge(x=dbh_melt,y=now_melt,by=c("case", "lon","lat","time","dbh"),all=TRUE) #this merge gets too slow when data is large,use                                                                                                  #left_join
        dbh_melt = dbh_melt %>% left_join(now_melt,by=c("case", "lon","lat","time","dbh"))
      }#end if (is.null(scls.melt))
}#end for (d in sequence(nbydbh))

}

bydbh = dbh_melt
rm(dbh_melt)

cat0(" + Turn PFT-dependent matrices into data tables.")
pft_melt = NULL
loop_p   = c("fates_basalarea","fates_crownarea","fates_vegc_aboveground","fates_lai","fates_gpp","fates_mortality_background","fates_mortality_cstarv","fates_mortality_freezing","fates_mortality_fire","fates_mortality_hydraulic","fates_mortality_termination","fates_nplant")
n_loop   = length(loop_p)
for (c in sequence(ncase)){
  case = case_names[c]
  now_pftmelt = NULL
  for (p in sequence(n_loop)){
  #f        = match(names(bypft[[case]])[p],fatesvar_new$vnam)
  p_now    = loop_p[p] 
  f        = match(p_now,fatesvar_new$vnam)
  f_vnam   = fatesvar_new$vnam[f]
  f_desc   = fatesvar_new$desc[f]
  cat0("   - ",f_desc,".")
  
  now_pft           = as_tibble(bypft[[case]][[p_now]])
  #now_pft           = as_tibble(bypft[[case]][[f_vnam]])
  now_pft$lon       = rep(lon_vec,times=nstep)
  now_pft$lat       = rep(lat_vec,times=nstep)
  now_pft$time      = ens_tstamp
  now_pft$case      = rep_along(time, x=case)
  now_melt          = as_tibble(reshape2::melt(data=now_pft,id.vars=c("case","lon","lat","time"),variable.name="pft",value.name=f_vnam))

  if (is.null(now_pftmelt)){
      now_pftmelt = now_melt
      }else{
         now_pftmelt = now_pftmelt %>% left_join(now_melt,by=c("case", "lon", "lat","time", "pft"))
      }
}#end for (p in sequence(nbypft))
  pft_melt = rbind(pft_melt,now_pftmelt)
}


bypft = pft_melt
rm(pft_melt)
invisible(gc())

#cat0(" + Turn fuel-class matrices into tibble objects.")
#fuel_melt  = NULL

#for (c in sequence(ncase)){
#  case = case_names[c]
#  now_fuelmelt = NULL
#  for (u in sequence(nbyfuel)){
   #--- Match variables.
#  f        = match(names(byfuel[[case]])[u],fatesvar$vnam)
#  f_vnam   = fatesvar$vnam[f]
#  f_desc   = fatesvar$desc[f]
#  cat0("   - ",f_desc,".")
      
#--- Create molten data table for this variable.  
#  now_fuel            = as_tibble(byfuel[[case]][[f_vnam]])
#  now_fuel$lon        = rep(lon_vec,time=nstep)
#  now_fuel$lat        = rep(lat_vec,time=nstep)
#  now_fuel$time       = ens_tstamp
#  now_fuel$case       = rep_along(time, x = case)
#  now_melt            = as_tibble(reshape2::melt(data=now_fuel,id.vars=c("case","lon",       
#                              "lat","time"),variable.name="fuel",value.name=f_vnam))
      #--- Merge data table
  
#  if (is.null(now_fuelmelt)){
#    now_fuelmelt = now_melt
#  }else{
#        now_fuelmelt = as_tibble(merge(x=now_fuelmelt,y=now_melt,by=c("case","lon","lat","time","fuel"),all=TRUE))
#  }#end if (is.null(age_melt))
  
 # }#end for (a in sequence(nbyage))
#fuel_melt = rbind(fuel_melt,now_fuelmelt)
#}



#byfuel = fuel_melt 
#rm(fuel_melt)
#invisible(gc())

# Turn drought-deciduous phenology variables into data tables
#if (! is_tibble(dphen)){
#   cat0(" + Turn drought-deciduous phenology matrices into data tables.")
#   dph_melt = NULL
#   for (c in sequence(ncase)){
#     case = case_names[c]
#     now_dphmelt = NULL
#     for (p in sequence(ndphen)){
      # Match variables.
#      f        = match(names(dphen[[case]])[p],dphenvar$vnam)
#      f_vnam   = dphenvar$vnam[f]
#      f_desc   = dphenvar$desc[f]
#      cat0("   - ",f_desc,".")

      # Create molten data table for this variable.  
#      now_dph           = as_tibble(dphen[[case]][[p]])
#      now_dph$lon       = lon_vec
#      now_dph$lat       = lat_vec
#      now_dph$time      = ens_tstamp
#      now_dph$case      = rep_along(time, x=case)
#      now_melt          = as_tibble(melt(data=now_dph,id.vars=c("case","lon","lat","time"),
#                                    variable.name="pft",value.name=f_vnam))

      # Merge data table
#      if (is.null(now_dphmelt)){
#         now_dphmelt = now_melt
#      }else{
#         now_dphmelt = as_tibble(merge(x=now_dphmelt,y=now_melt,by=c("case","lon","lat","time","pft"),all=TRUE))
#      }#end if (is.null(dph_melt))
#   }#end for (p in sequence(ndphen))
#    dph_melt = rbind(dph_melt,now_dphmelt)
#}# end of for (c insequence(ncase))
   
   # Replace dphen with the tibble object
#   dphen = dph_melt
#   rm(dph_melt)
#   invisible(gc())
#}#end if (! is_tibble(dphen))

#cat0(" + Turn soil layer-dependent matrices into data tables.")
#soi_melt = NULL
#for (c in sequence(ncase)){
#  case = case_names[c]
#  now_soimelt = NULL
#  for (s in sequence(nhlm2d)){
#  h        = match(names(hlm2d[[case]])[s],hlm2dsoi$vnam)
#  h_vnam   = hlm2dsoi$vnam[h]
#  h_desc   = hlm2dsoi$desc[h]
#  cat0("   - ",h_desc,".")
   
#  now_soi           = as_tibble(hlm2d[[case]][[s]])
#  now_soi$lon       = lon_vec
#  now_soi$lat       = lat_vec
#  now_soi$time      = ens_tstamp
#  now_soi$case      = rep_along(time, x=case)
#  now_melt     = as_tibble(reshape2::melt(data=now_soi,id.vars=c("case", "lon",                 
#                                             "lat","time"),variable.name="soi",value.name=h_vnam))
#  if (is.null(now_soimelt)){
#      now_soimelt = now_melt
#      }else{
#         now_soimelt = merge(x=now_soimelt,y=now_melt,by=c("case", "lon","lat","time","soi"),all=TRUE)
#      }#end if (is.null(scls.melt))
#  }#end for (s in sequence(nsoi))
#  soi_melt = rbind(soi_melt,now_soimelt)
  
#}


#bysoi = soi_melt
#rm(soi_melt)
#invisible(gc())


cat0(" + Convert classes to integers.")
#byage$age = as.integer(byage$age)
bydbh$dbh = as.integer(bydbh$dbh)
bypft$pft   = as.integer(bypft$pft)
#byfuel$fuel = as.integer(byfuel$fuel)
#dphen$pft = as.integer(dphen$pft)
#bysoi$soi = as.integer(bysoi$soi)

## write out tables so can use later
fwrite(bypft,file.path(main_path,paste0(file_desc,"_bypft.csv")))
fwrite(bydbh,file.path(main_path,paste0(file_desc,"_bydbh.csv")))


bypft  = fread(file.path(main_path,paste0(file_desc,"_bypft.csv")))
hlm1d  = fread(file.path(main_path,paste0(file_desc,"_hlm1d.csv")))
bydbh = fread(file.path(main_path,paste0(file_desc,"_bydbh.csv")))

```









```{r,label="load-observations",message=FALSE,results='hide'}
cat0(" + Load regional data from ",obs_main,".")

growing_sea = c(2,3,4,5,6)
fire_month    = c(8,9,10,11)

## load FRAP regional burned fraction range

frap_bf    = fread(frap_file)
frap_bf    =  frap_bf                              %>% 
              filter(year %in% simu_time)          %>% 
              mutate(lon=round(lon,digits=5)
                   , lat=round(lat,digits=5))      

frap_bf$lon_f = factor(frap_bf$lon,levels=unique(frap_bf$lon))
frap_bf$lat_f = factor(frap_bf$lat,levels=unique(frap_bf$lat))

   
frap_amean = frap_bf                                   %>%
             group_by(lon_f,lat_f)                     %>% 
             mutate(bfrac=mean(bfrac,na.rm=TRUE))      %>% 
             ungroup()                                  
             

frap_qupr  = frap_bf                                   %>% 
             group_by(lon_f,lat_f)                     %>% 
             summarize(bfrac_qupr=quantile(bfrac
                                          ,probs=qupr_ribbon
                                          ,na.rm=TRUE)) %>% 
             ungroup()                                 

frap_qlwr  = frap_bf                                   %>% 
             group_by(lon_f,lat_f)                     %>% 
             summarize(bfrac_qlwr=quantile(bfrac
                                          ,probs=qlwr_ribbon
                                          ,na.rm=TRUE)) %>% 
             ungroup()

frap_amean = frap_amean                                  %>% 
             left_join(frap_qupr,by=c("lon_f","lat_f"))  %>% 
             left_join(frap_qlwr,by=c("lon_f","lat_f"))  %>% 
             dplyr::select(-c(lon_f,lat_f,year))         %>% 
             distinct()




## range of annual burned fraction by each year using regional data         
bf_quantyr = frap_bf                                 %>% 
             group_by(year)                          %>%
             summarize(yr_qupr = quantile(bfrac,probs=qupr_ribbon,na.rm=TRUE)
                      ,yr_qlwr = quantile(bfrac,probs=qlwr_ribbon,na.rm=TRUE)
                      ,yr_mean = mean(bfrac,na.rm=TRUE)) %>% 
             ungroup()



### load gpp, lai, agb, basal area, and canopy cover and merge them 
gpp_ob = fread(gpp_path)
lai_ob = fread(lai_path)
agb_ob = fread(mass_path)
ba_ob  = fread(bsarea_path)
cv_ob  = fread(cancov_path)
obs_name = c("gpp_ob","lai_ob","agb_ob","ba_ob","cv_ob")
obs_desc = c("gpp","lai","biomass","oakBA","cancov")
n_obs    = length(obs_desc)

all_obs   = NULL
grow_obs  = NULL

for(v in sequence(n_obs)){
    obn    = obs_name[v]
    obd    = obs_desc[v]
    ob_now = eval(parse(text = obn))
    ob_now = ob_now                                 %>% 
             mutate(lon = round(lon,digits=5),
                    lat = round(lat,digits=5))
    ob_now$lon_f = factor(ob_now$lon,levels=unique(ob_now$lon))
    ob_now$lat_f = factor(ob_now$lat,levels=unique(ob_now$lat))
    orig_xy      = ob_now %>% dplyr::select(lon,lat,lon_f,lat_f) %>% distinct()
    if(obn %in% c("gpp_ob","lai_ob")){
      ob_ameanAll  = ob_now                                 %>% 
                     dplyr::select(-c(lon,lat,year,month))  %>% 
                     group_by(lon_f,lat_f)                  %>% 
                     summarize_all(mean)                    %>% 
                     ungroup()
      ob_ameanAll = orig_xy %>% left_join(ob_ameanAll,by=c("lon_f","lat_f"))
      ob_ameanGrow = ob_now                                  %>% 
                     filter(month%in%growing_sea)            %>% 
                     dplyr::select(-c(lon,lat,year,month))   %>% 
                     group_by(lon_f,lat_f)                   %>% 
                     summarize_all(mean)                     %>% 
                     ungroup()
     ob_ameanGrow = orig_xy %>% left_join(ob_ameanGrow,by=c("lon_f","lat_f"))
    }else{
       ob_ameanAll  = ob_now                                 %>% 
                      dplyr::select(-c(lon,lat))             %>% 
                      group_by(lon_f,lat_f)                  %>% 
                      summarize_all(mean)                    %>% 
                      ungroup()
       ob_ameanAll = orig_xy %>% left_join(ob_ameanAll,by=c("lon_f","lat_f"))
    }
    
    
  if (is.null(all_obs)){
     all_obs = ob_ameanAll
    }else{all_obs = merge(x=all_obs,y=ob_ameanAll,by=c("lon_f","lat_f","lon","lat"),all=TRUE)}
    
  if(is.null(grow_obs) & obn %in% c("gpp_ob","lai_ob")){
      grow_obs = ob_ameanGrow
    }else if(!is.null(grow_obs) & obn %in% c("gpp_ob","lai_ob")){
      grow_obs = merge(x=grow_obs,y=ob_ameanGrow,by=c("lon_f","lat_f","lon","lat"),all=TRUE)
    }
  }

## join GPP, and LAI
ras_ameanAll  = all_obs
ras_ameanGrow = grow_obs  
rm(lai_ob,gpp_ob,agb_ob,cv_ob,ba_ob,all_obs,grow_obs)

## total burned fraction by year from FRAP, SHORT, and a merged FRAP-SHORT(only fire <300 acer)
#tbf    = fread(tbf_grass)
#tbf    = tbf %>% filter(year %in% simu_time)


```









```{r,label="fates-variable-mean",fig.height=50, fig.width=40,message=FALSE}

mortout    = names(bypft)[grep(pattern="mortality",x=names(bypft))]

find_mort = function(x,n){
    mort_mon = ifelse( test = n == 0, yes = 0., no = pmin(1.,x/n/12.))
    mort_year = 100. * (1. - (1. - mort_mon)^12)
    return(mort_year)
}#end function find_mort


all    =     bypft                                             %>%  
             mutate(month = month(time)
                   ,year  = year(time))                        %>% 
             mutate_at(all_of(mortout)
                      ,~ find_mort(x=.x,n=.data$fates_nplant)) %>% 
             mutate(lon = round(lon,5)
                   ,lat = round(lat,5))
  
all$lon_f = factor(all$lon,levels=unique(all$lon))
all$lat_f = factor(all$lat,levels=unique(all$lat))

orig_xy   = hlm1d  %>% dplyr::select(lon,lat) %>% distinct() %>% 
            mutate(lon=round(lon,5)
                  ,lat=round(lat,5)
                  ,lon_f = factor(lon,levels=unique(lon))
                  ,lat_f = factor(lat,levels=unique(lat)))

cell_mmean = all                                                     %>% 
            group_by(case,lon_f,lat_f,pft,month)                     %>%  
  summarize( fates_mortality_background   = mean(fates_mortality_background)
           , fates_mortality_cstarv       = mean(fates_mortality_cstarv)
           , fates_mortality_freezing     = mean(fates_mortality_freezing)
           , fates_mortality_hydraulic    = mean(fates_mortality_hydraulic)
           , fates_mortality_termination  = mean(fates_mortality_termination)
           , fates_mortality_fire         = mean(fates_mortality_fire)
           , fates_vegc_aboveground       = mean(fates_vegc_aboveground)
           , fates_crownarea              = mean(fates_crownarea)
           , fates_basalarea              = mean(fates_basalarea)
           , fates_lai                    = mean(fates_lai)
           , fates_gpp                    = mean(fates_gpp)
           )  %>% 
  ungroup()                                                                   

cell_mmean = orig_xy %>% left_join(cell_mmean,by=c("lon_f","lat_f")) 


cell_amean =  all                                                        %>% 
              group_by(case,lon_f,lat_f,pft,year)                        %>% 
              summarize(fates_mortality_background   = mean(fates_mortality_background)
                      , fates_mortality_cstarv       = mean(fates_mortality_cstarv)
                      , fates_mortality_freezing     = mean(fates_mortality_freezing)
                      , fates_mortality_hydraulic    = mean(fates_mortality_hydraulic)
                      , fates_mortality_termination  = mean(fates_mortality_termination)
                      , fates_mortality_fire         = mean(fates_mortality_fire)
                      , fates_vegc_aboveground       = mean(fates_vegc_aboveground)
                      , fates_crownarea              = mean(fates_crownarea)
                      , fates_basalarea              = mean(fates_basalarea)
                      , fates_lai                    = mean(fates_lai)
                      , fates_gpp                    = mean(fates_gpp))   %>%
               ungroup()                                                 
 
cell_amean = orig_xy %>% left_join(cell_amean,by=c("lon_f","lat_f"))
 
 
 cellmean_struc  = all                                                                   %>% 
              mutate(year = year(time))                                                  %>% 
              filter(year %in% c(1977:1981))                                             %>% 
              group_by(case,lon_f,lat_f,pft)                                             %>% 
              summarize(#fates_mortality_background   = mean(fates_mortality_background)
                      #, fates_mortality_cstarv       = mean(fates_mortality_cstarv)
                      #, fates_mortality_freezing     = mean(fates_mortality_freezing)
                      #, fates_mortality_hydraulic    = mean(fates_mortality_hydraulic)
                      #, fates_mortality_termination  = mean(fates_mortality_termination)
                      #, fates_mortality_fire         = mean(fates_mortality_fire)
                      #, 
                    fates_vegc_aboveground       = mean(fates_vegc_aboveground)
                      , fates_crownarea              = mean(fates_crownarea)
                      , fates_basalarea              = mean(fates_basalarea)
                      , fates_lai                    = mean(fates_lai)
                      , fates_gpp                    = mean(fates_gpp))             %>%
              ungroup()   
 
 
cellmean_struc   = orig_xy  %>% left_join(cellmean_struc,by=c("lon_f","lat_f"))
 
 cellmean_mort    = all %>% #filter(year %in% c(1:1885))  %>% 
                    group_by(case,lon_f,lat_f,pft) %>% 
                    summarize(fates_mortality_background   = mean(fates_mortality_background), 
                              fates_mortality_cstarv       = mean(fates_mortality_cstarv), 
                              fates_mortality_freezing     = mean(fates_mortality_freezing), 
                              fates_mortality_hydraulic    = mean(fates_mortality_hydraulic), 
                              fates_mortality_termination  = mean(fates_mortality_termination), 
                              fates_mortality_fire         = mean(fates_mortality_fire)) %>% ungroup()
 
cellmean_mort   = orig_xy  %>% left_join(cellmean_mort,by=c("lon_f","lat_f"))


  
```






```{r,label="plot-pft-mean",message=FALSE,results="hide"}

plot_mort  =  c("fates_mortality_background"
              ,"fates_mortality_cstarv" 
              , "fates_mortality_freezing"
              , "fates_mortality_fire"
              , "fates_mortality_hydraulic"
              , "fates_mortality_termination")
plot_struc = c(
               "fates_vegc_aboveground"
              , "fates_crownarea"
              , "fates_basalarea"
              , "fates_gpp"
              , "fates_lai")
ca_co     = USAboundaries::us_counties(resolution = "high", states = "CA")
wgs = "+init=EPSG:4326"
mort_loop = which(fatesvar_new$vnam %in% plot_mort)
struc_loop = which(fatesvar_new$vnam %in% plot_struc)

gg_emean   = list()
for(c in sequence(ncase)){
  case_now  = case_names[c]
  desc_now  = case_desc[c]
  mmean_now = cellmean_struc %>% filter(case == case_now) 
for(p in sequence(npfts)){
  pft_now = pftinfo$id[p]
  pft_desc=pftinfo$short[p]
  var_df  =mmean_now %>% filter(pft == pft_now)
  for (h in struc_loop){
   f_vnam   = fatesvar_new$vnam[h] 
   f_desc   = fatesvar_new$desc[h]
   f_unit   = fatesvar_new$unit[h]
   f_vqlwr  = paste0(f_vnam,"_qlwr")
   f_vqupr  = paste0(f_vnam,"_qupr")
   h_legend =  1
   cat0("   - ",f_desc,".")
   plot_df   = var_df %>% dplyr::select(all_of(c("lon","lat",f_vnam)))
   
   
   fil_var  = matrix(plot_df[[f_vnam]], nrow=147,ncol=151)
   x_arr    = matrix(plot_df$lon,nrow=147,ncol=151)
   y_arr    = matrix(plot_df$lat,nrow=147,ncol=151)
   wrf_star = st_as_stars(fil_var)
   wrf_star = st_as_stars(wrf_star, curvilinear=list(X1=x_arr,X2=y_arr), crs=wgs)
   wrf_sf   = st_as_sf(wrf_star,as_points=FALSE,na.rm=FALSE)
   
   gg_now = ggplot()
   gg_now = gg_now + geom_sf(data=wrf_sf,aes(fill=A1),color=NA)
   gg_now = gg_now + coord_sf(crs=st_crs(wgs)) 
   gg_now = gg_now + scale_fill_gradientn(colours=cols, guide="colorbar", na.value="white")
   gg_now = gg_now + geom_sf(data = ca_co, color = alpha("black", alpha=1),lwd=0.1,fill=NA) 
   gg_now = gg_now + scale_x_continuous(limits = c(-125, -112),expand = c(0,0))
   gg_now = gg_now + scale_y_continuous(limits = c(32, 42.5),expand = c(0,0))
   gg_now = gg_now + labs(title=desc_now
                         , x="Longitude"
                         , y="Latitude"
                         , fill=desc.unit(desc=f_desc,unit=untab[[f_unit]],twolines =TRUE))
   gg_now = gg_now + theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size= 0.5,base_rect_size =0.5)
   gg_now = gg_now + theme( legend.position   = "right" 
                          , panel.background  = element_blank()
                          , panel.border      = element_rect(linewidth = 1.6, fill=NA)
                          , axis.text.x       = element_text( size   = gg_ptsz-5
                                                           , margin = unit(rep(0.35,times=4),"cm"))#end element_text
                          , axis.text.y       = element_text( size   = gg_ptsz-5
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                          , axis.ticks.length = unit(-0.25,"cm")
                         ) #end theme
  
 
   # Save plots.
   for (d in sequence(ndevice)){
    h_output = paste0(case_now,f_vnam,pft_desc,"end.",gg_device[d])
    dummy    = ggsave( filename = h_output
                    , plot     = gg_now
                    , device   = gg_device[d]
                    , path     = secy_path
                    , width    = gg_widthn*0.6
                    , height   = gg_height*0.6
                    , units    = gg_units
                    , dpi      = gg_depth
                    )
   }
  gg_emean[[f_vnam]] = gg_now      

    }
  }
}


# If sought, plot images on screen
if (gg_screen) gg_emean



```








```{r,label="hlm-variable-summary-sets",results='hide'}

hlm_var   = c("fates_gpp","fates_vegc_aboveground",
              "eflx_lh_tot","elai","fsh")
hlm1d       = hlm1d                       %>% 
              mutate( lon = round(lon,5)
                    , lat = round(lat,5))
hlm1d$lon_f = factor(hlm1d$lon,levels=unique(hlm1d$lon))
hlm1d$lat_f = factor(hlm1d$lat,levels=unique(hlm1d$lat))
orig_xy     = hlm1d %>% dplyr::select(case,lon,lat,lon_f,lat_f) %>% distinct()

bf_mod   =  hlm1d                                                                  %>% 
            dplyr::select(all_of(c("case","time","lon","lat",
                                   "lon_f","lat_f","fates_burnfrac",
                                   "fates_fuel_amount","rain",
                                   "q2m","qaf","taf","tsa","u10")))                %>% 
            mutate(month=month(time))                                              %>% 
            filter(month %in% fire_month)                                          %>% 
            dplyr::select(-month)


hlm_scy    = hlm1d                                                                %>% 
             dplyr::select(all_of(c("case","time","lon","lat",
                                    "lon_f","lat_f", hlm_var,
                                    "rain","tsa","taf","q2m")))                   %>% 
             mutate(month=month(time),
                    year=year(time))                                              %>% 
             filter(month%in%growing_sea & year%in%c(2010:2014))                  %>% 
             dplyr::select(-c(month,year))



mean_bf =   bf_mod                                     %>%
            group_by(case,lon_f,lat_f)                 %>%
            summarize(fates_burnfrac=mean(fates_burnfrac)
                     ,fates_fuel_amount=mean(fates_fuel_amount)
                     ,q2m_d = mean(q2m)
                     ,qaf_d = mean(qaf)
                     ,taf_d = mean(taf)
                     ,tsa_d = mean(tsa)
                     ,u10   = mean(u10)
                     )                                 %>% 
            ungroup()   
mean_bf  = orig_xy %>% left_join(mean_bf,by=c("case","lon_f","lat_f")) %>% dplyr::select(-c(lon_f,lat_f))


mean_bfall = hlm1d                                          %>% 
             group_by(case,lon_f,lat_f)                     %>% 
             summarize(fates_burnfrac=mean(fates_burnfrac)
                      ,fates_fuel_amount=mean(fates_fuel_amount)
                      ,qaf  = mean(qaf)
                      ,u10 = mean(u10))                     %>% 
             ungroup()                                      
mean_bfall = orig_xy %>% left_join(mean_bfall,by=c("case","lon_f","lat_f")) %>% dplyr::select(-c(lon_f,lat_f))                                       

sum_bfall  = hlm1d                                     %>% 
             group_by(case,lon_f,lat_f)                %>% 
             mutate(bfrac=sum(fates_burnfrac))         %>% 
             ungroup()                                 %>% 
             dplyr::select(case,lon,lat,bfrac)         %>% 
             distinct()                                %>% 
             mutate(bfrac=bfrac*yr.day) 


mean_hlm   = hlm_scy                          %>% 
             group_by(case,lon_f,lat_f)       %>%
             summarize(fates_gpp = mean(fates_gpp)
                      ,fates_vegc_aboveground = mean(fates_vegc_aboveground)
                      ,eflx_lh_tot = mean(eflx_lh_tot)
                      ,elai = mean(elai)
                      ,fsh  = mean(fsh)
                      ,rain = mean(rain)
                      ,tsa_g  = mean(tsa)
                      ,taf_g  = mean(taf)
                      ,q2m_g  = mean(q2m))    %>% 
             ungroup()
mean_hlm = orig_xy %>% left_join(mean_hlm,by=c("case","lon_f","lat_f")) %>% dplyr::select(-c(lon_f,lat_f))

allmean_hlm = hlm1d                                      %>%
              group_by(case,lon_f,lat_f)                 %>% 
              summarize(fates_gpp = mean(fates_gpp)
                       ,fates_vegc_aboveground = mean(fates_vegc_aboveground)
                       ,eflx_lh_tot = mean(eflx_lh_tot)
                       ,elai = mean(elai)
                       ,fsh = mean(fsh)
                       ,rain = mean(rain)
                       ,tsa  = mean(tsa)
                       ,taf  = mean(taf)
                       ,q2m  = mean(q2m))                 %>% 
               ungroup()


allmean_hlm = orig_xy %>% left_join(allmean_hlm,by=c("case","lon_f","lat_f")) %>% dplyr::select(-c(lon_f,lat_f))
## mean of each variable from either peak growing season or fire season
mean_pseason = mean_hlm %>% left_join(mean_bf,by=c("case","lon","lat")) %>% ungroup()

## mean of each variable from all months
mean_allseson = allmean_hlm  %>% left_join(mean_bfall,by=c("case","lon","lat")) %>% ungroup()

## merge FRAP mean burned fraction and GPP, LAI observations as well
grosea_hlmcom = mean_pseason                               %>% 
                mutate(lon = round(lon,5),
                       lat = round(lat,5))                 %>% 
                left_join(ras_ameanGrow,by=c("lon","lat")) %>% 
                left_join(frap_amean,by=c("lon","lat"))


allsea_hlmcom = mean_allseson                              %>% 
                mutate(lon = round(lon,5)
                      ,lat = round(lat,5))                 %>% 
                left_join(ras_ameanAll,by=c("lon","lat"))  %>% 
                left_join(frap_amean,by=c("lon","lat"))

#filter obs to only where land cell is active
grosea_hlmcom =  grosea_hlmcom        %>% 
                 mutate(gpp = ifelse(!is.na(fates_gpp), gpp, NA)
                       ,lai = ifelse(!is.na(elai),lai,NA)
                       ,bfrac = ifelse(!is.na(fates_burnfrac),bfrac,NA))

allsea_hlmcom =  allsea_hlmcom        %>% 
                 mutate(gpp = ifelse(!is.na(fates_gpp), gpp, NA)
                       ,lai = ifelse(!is.na(elai),lai,NA)
                       ,biomass=ifelse(!is.na(fates_vegc_aboveground),biomass,NA)
                       ,bfrac = ifelse(!is.na(fates_burnfrac),bfrac,NA)
                       ,oakBA = ifelse(!is.na(fates_vegc_aboveground),oakBA,NA)
                       ,cancov=ifelse(!is.na(fates_vegc_aboveground),cancov,NA))


## summed burned fraction by year
bf_simu = mean_bf %>% dplyr::select(case,lon,lat,fates_burnfrac) %>% 
          mutate(lon=round(lon,digits=5)
                ,lat=round(lat,digits=5))     
bfsum_simu = sum_bfall %>% dplyr::select(case,lon,lat,bfrac) %>% 
             mutate(lon=round(lon,digits=5)
                   ,lat=round(lat,digits=5))

#bfsum_ob   = cags_bf %>% group_by(lon_f,lat_f)   %>% 
#             mutate(bfrac_ob=sum(bfrac,na.rm=TRUE)) %>% 
#             ungroup()                           %>% 
#             dplyr::select(lon,lat,bfrac_ob)     %>% 
#             distinct()   
#bfall_comm = bfsum_simu %>% left_join(bfsum_ob,by=c("lon","lat"))

## relative burned fraction

#bf_comm = bf_simu  %>% left_join(cags_dmean,by=c("lon","lat")) %>% 
#          mutate(rela_bf = fates_burnfrac/bfrac)               %>% 
#          mutate(rela_bf = ifelse(is.infinite(rela_bf),NA,rela_bf))

```













```{r,label="plot-HLM-mean"}
cat0(" + Plot  mean of HLM variables.")

#leg_colours = c(site_colour,ens_colour)
#leg_labels  = c("Site",case_desc)
plot_hlm  =  c("fates_gpp","fates_fuel_amount","fates_vegc_aboveground","elai","fates_burnfrac",
               "tsa","q2m","rain","u10")
ca_co     = USAboundaries::us_counties(resolution = "high", states = "CA")
wgs = "+init=EPSG:4326"
emean_loop = which(hlm1dvar$vnam %in% plot_hlm)

gg_emean   = list()
for(c in sequence(ncase)){
  case_now = case_names[c]
  desc_now = case_desc[c]
mmean_now        = mean_allseson %>% filter(case == case_now) 
for (h in emean_loop){
   h_vnam   = hlm1dvar$vnam[h] 
   h_desc   = hlm1dvar$desc [h]
   h_short  = hlm1dvar$short[h]
   h_unit   = hlm1dvar$unit [h]
   h_vqlwr  = paste0(h_vnam,"_qlwr")
   h_vqupr  = paste0(h_vnam,"_qupr")
   h_legend =  1
   cat0("   - ",h_desc,".")
   var_df   = mmean_now %>% dplyr::select(all_of(c("lon","lat",h_vnam)))
   
   
   fil_var  = matrix(var_df[[h_vnam]], nrow=147,ncol=151)
   x_arr    = matrix(var_df$lon,nrow=147,ncol=151)
   y_arr    = matrix(var_df$lat,nrow=147,ncol=151)
   wrf_star = st_as_stars(fil_var)
   wrf_star = st_as_stars(wrf_star, curvilinear=list(X1=x_arr,X2=y_arr), crs=wgs)
   wrf_sf   = st_as_sf(wrf_star,as_points=FALSE,na.rm=FALSE)
   
   gg_now = ggplot()
   gg_now = gg_now + geom_sf(data=wrf_sf,aes(fill=A1),color=NA)
   gg_now = gg_now + coord_sf(crs=st_crs(wgs)) 
   gg_now = gg_now + scale_fill_gradientn(colours=cols, guide="colorbar", na.value="white")
   gg_now = gg_now + geom_sf(data = ca_co, color = alpha("black", alpha=1),lwd=0.1,fill=NA) 
   gg_now = gg_now + scale_x_continuous(limits = c(-125, -112),expand = c(0,0))
   gg_now = gg_now + scale_y_continuous(limits = c(32, 42.5),expand = c(0,0))
   gg_now = gg_now + labs(title=desc_now
                         , x="Longitude"
                         , y="Latitude"
                         , fill=desc.unit(desc=h_short,unit=untab[[h_unit]],dxpr=TRUE))
   gg_now = gg_now + theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size= 0.5,base_rect_size =0.5)
   gg_now = gg_now + theme( legend.position   = "right" 
                          , panel.background  = element_blank()
                          , panel.border      = element_rect(linewidth = 1.6, fill=NA)
                          , axis.text.x       = element_text( size   = gg_ptsz-5
                                                           , margin = unit(rep(0.35,times=4),"cm"))#end element_text
                          , axis.text.y       = element_text( size   = gg_ptsz-5
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                          , axis.ticks.length = unit(-0.25,"cm")
                         ) #end theme
  
 
   # Save plots.
   for (d in sequence(ndevice)){
    h_output = paste0(case_now,h_vnam,"Transallmean.",gg_device[d])
    dummy    = ggsave( filename = h_output
                    , plot     = gg_now
                    , device   = gg_device[d]
                    , path     = secy_path
                    , width    = gg_widthn*0.6
                    , height   = gg_height*0.6
                    , units    = gg_units
                    , dpi      = gg_depth
                    )
   }
  gg_emean[[h_vnam]] = gg_now      

}
}


# If sought, plot images on screen
if (gg_screen) gg_emean

```







```{r,label="model-to-observation-spatial-and-scatter-plot"}

cat0(" + Plot  model observation comparison.")      

plot_vars  = c("fates_gpp","elai","fates_burnfrac","fates_vegc_aboveground")
match_obs  = c("gpp","lai","bfrac","biomass")
obs_label  = c("SIF","MODIS","FRAP","LEMMA")
#cols <- brewer.pal(n = 5, name = "RdBu",direction=-1) 

nloop=length(plot_vars)
for(n in sequence(ncase)){
  all_plots = list()
  case_now = case_names[n]
  desc_now = case_desc[n]
  bf_now  = allsea_hlmcom %>% filter(case==case_now)
  for (p in sequence(nloop)){
    
    var_now = plot_vars[p]
    ob_var = match_obs[p]
    ob_label = obs_label[p]
    h = which(hlm1dvar$vnam %in% var_now)
    h_vnam   = hlm1dvar$vnam[h] 
    h_desc   = hlm1dvar$desc [h]
    h_short  = hlm1dvar$short[h]
    h_unit   = hlm1dvar$unit [h]
    

  
  
  #if(var_now!="fates_burnfrac"){
   # ylim   = range(c(bf_now[[var_now]], bf_now[[ob_var]]),na.rm=TRUE) #}else{
    y = c(bf_now[[var_now]], bf_now[[ob_var]])
    rng = range(y,na.rm=TRUE)
    qn     = quantile(y, c(0.25,0.95),na.rm=TRUE)
    qn_rescle = rescale(c(qn, range(y,na.rm=TRUE)))
    brks      = c(0, seq(qn_rescle[1],qn_rescle[2],length.out=3),1)
    
   #}
  
    fil_mod  = matrix(bf_now[[var_now]], nrow=147,ncol=151)
    x_arr    = matrix(bf_now$lon,nrow=147,ncol=151)
    y_arr    = matrix(bf_now$lat,nrow=147,ncol=151)
    mod_star = st_as_stars(fil_mod)
    mod_star = st_as_stars(mod_star, curvilinear=list(X1=x_arr,X2=y_arr), crs=wgs)
    mod_sf   = st_as_sf(mod_star,as_points=FALSE,na.rm=FALSE)
    wd_ratio = tmaptools::get_asp_ratio(mod_sf)
    ca_co    = USAboundaries::us_counties(resolution = "high", states = "CA")

    mod = ggplot() + geom_sf(data=mod_sf, aes(fill=A1),color=NA)+
    coord_sf(crs=st_crs(wgs)) + 
    scale_fill_gradientn(colours=cols, guide="colorbar", na.value="white", 
                      values=brks, limits=rng,oob = scales::squish) +
  #scale_fill_continuous(low="blue", high="red", na.value="white", guide="colorbar", limits=ylim,oob = scales::squish)+
    geom_sf(data = ca_co, color = alpha("black", alpha=1),lwd=0.1,fill=NA) +
    scale_x_continuous(limits = c(-125, -112),expand=c(0,0)) +
    scale_y_continuous(limits = c(32, 42.5),expand=c(0,0)) + 
    labs(x=NULL,y=NULL,title="FATES") +
    guides(fill=guide_legend(title=desc.unit(desc=h_short,unit=untab[[h_unit]],dxpr=TRUE))) +
    theme_grey( base_size = gg_ptsz*0.8, base_family = "Helvetica",base_line_size= 0.5,base_rect_size =0.5) +
    theme(  legend.position  = "right"
        , legend.title     = element_text(size=gg_ptsz*0.8)
        , legend.text      = element_text(size=gg_ptsz*0.8)
        , plot.title       = element_text(size=gg_ptsz*0.8, vjust=1)
        , panel.background = element_blank()
        #, panel.border     = element_rect(linewidth = 1.6, fill=NA)
        , panel.border     = element_blank()
        #, axis.text.x      = element_text( size   = gg_ptsz*0.85
        , axis.text.x      = element_blank()
        #, margin           = unit(rep(0.35,times=4),"cm")#end element_text
        #, axis.text.y      = element_text( size   = gg_ptsz*0.85
        , axis.text.y      = element_blank()
        #, margin           = unit(rep(0.35,times=4),"cm")
        , axis.ticks=element_blank()
        ,plot.margin = unit(c(0,0,0,0), "cm")
      )#end element_text
     #, axis.ticks.length = unit(-0.25,"cm")
#      ) #end theme

  fil_obs  = matrix(bf_now[[ob_var]], nrow=147,ncol=151)
  obs_star = st_as_stars(fil_obs)
  obs_star = st_as_stars(obs_star, curvilinear=list(X1=x_arr,X2=y_arr), crs=wgs)
  obs_sf   = st_as_sf(obs_star,as_points=FALSE,na.rm=FALSE)

  obs = ggplot() + geom_sf(data=obs_sf, aes(fill=A1),color=NA)+
  coord_sf(crs=st_crs(wgs)) + 
  #scale_fill_continuous(low="blue", high="red", na.value="white", guide="colorbar", limits=ylim,oob = scales::squish)+
  scale_fill_gradientn(colours=cols, guide="colorbar", na.value="white",
                       values = brks, limits=rng,oob=scales::squish) +
  geom_sf(data = ca_co, color = alpha("black", alpha=1),lwd=0.1,fill=NA) +
  scale_x_continuous(limits = c(-125, -112),expand=c(0,0)) +
  scale_y_continuous(limits = c(32, 42.5),expand=c(0,0)) +
  labs(x=NULL,y=NULL,title=ob_label) +
  guides(fill=guide_legend(title=desc.unit(desc=h_short,unit=untab[[h_unit]],dxpr=TRUE))) +
  #guides(fill=guide_legend(title=desc.unit(desc=paste0(ob_label,"-",h_short),unit=untab[[h_unit]],dxpr=TRUE))) +
  theme_grey( base_size = gg_ptsz*0.8, base_family = "Helvetica",base_line_size= 0.5,base_rect_size =0.5) +
  theme( legend.position   = "none"
        , legend.title     = element_text(size=gg_ptsz*0.8)
        , legend.text      = element_text(size=gg_ptsz*0.8)
        , panel.background = element_blank()
        , plot.title       = element_text(size=gg_ptsz*0.8, vjust=1)
        , panel.border = element_blank()
        , axis.text.x  = element_blank()
        , axis.text.y  = element_blank()
        , axis.ticks=element_blank()
        , plot.margin = unit(c(0,0,0,0), "cm")
      )#end element_text
     #, axis.ticks.length = unit(-0.25,"cm")
#      ) #end theme
 
  scat = ggplot() + geom_point(data=bf_now,aes_string(x=var_now,y=ob_var)) +
         labs(x="Model",y="Observation") +
         geom_abline(slope=1,intercept=0,linetype="dashed",color="black",linewidth=1) +
         theme_grey( base_size = gg_ptsz*0.8, base_family = "Helvetica",base_line_size= 0.5,base_rect_size =0.5)+
         theme( legend.position  = "none" 
              , panel.background = element_blank()
              , panel.border     = element_rect(linewidth = 1.6, fill=NA)
              , axis.text.x      = element_text( size   = gg_ptsz*0.8
              , margin           = unit(rep(0.35,times=4),"cm"))#end element_text
              , axis.text.y      = element_text( size   = gg_ptsz*0.8
                                               , margin = unit(rep(0.35,times=4),"cm")
                                               )#end element_text
              , axis.ticks.length = unit(-0.25,"cm")
              , plot.margin = unit(c(0,0,0,0), "cm")
              ) #end theme
  p1 = mod + obs +
     plot_layout(guides="collect",width = c(2,2,1), nrow=1) 
  p2 = p1 + scat
  #p = wrap_plots(mod,obs,scat) + 
     #plot_layout(guides="collect",width=c(2,2,1),ncol=3) 
   all_plots[[p]] = p2
   fname = paste0(case_now,"-",var_now,"_model-obs.png")
   ggsave(p2, filename=fname,path=secy_path,width=gg_widthn*0.4,height=gg_height*0.5,dpi=gg_depth,device=gg_device[1]) 
   fname = paste0(case_now,"-",var_now,"_mod.png")
   ggsave(mod, filename=fname,path=secy_path,width=gg_widthn*0.7,height=gg_height*0.7,dpi=gg_depth,device=gg_device[1])
   fname = paste0(case_now,"-",var_now,"_obs.png")
   ggsave(obs, filename=fname,path=secy_path,width=gg_widthn*0.7,height=gg_height*0.7,dpi=gg_depth,device=gg_device[1])
   
  }
  gg_patch = all_plots[[1]] / all_plots[[2]] / all_plots[3] / all_plots[[4]]
  for (d in sequence(ndevice)){
    h_output = paste0(case_now,"-mod-obs-all.",gg_device[d])
    dummy    = ggsave( filename = h_output
                    , plot      = gg_patch
                    , device    = gg_device[d]
                    , path      = secy_path
                    , width     = gg_widthn
                    , height    = gg_widthn*wd_ratio
                    , units     = gg_units
                    , dpi       = gg_depth
                    )
   }
}

 


### annual total burned fraction compare to FRAP

mo_allamean = hlm1d                                                        %>% 
              mutate(year=year(time))                                      %>% 
              group_by(case,year)                                          %>% 
              summarize(bf_amean=sum(fates_burnfrac,na.rm=TRUE))           %>% 
              mutate(bf_amean = bf_amean/749)                              %>%
              ungroup()                                                    %>% 
              mutate(case=case_desc[case])  %>% filter(case == "Quant-GENL-G1-N08T27")

ob_amean = tbf                                                             %>%
           dplyr::select(source,year,bfrac)                                %>% 
           rename( case=source, bf_amean=bfrac)

amean_comm = rbind(mo_allamean,ob_amean)
#amean_comm$case=factor(amean_comm$case,levels=unique(amean_comm$case))
gg_now = ggplot(amean_comm,aes(year,bf_amean,colour=case)) + geom_line() +
scale_color_manual(name="",values=top12) +
  labs(x="",y=expression("Summed burned fraction"~"("*yr^-1*")"))+
  theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size= 0.5,base_rect_size =0.5)+
  theme( legend.position   = "right" 
       , legend.text = element_text(size=gg_ptsz-5)
       , panel.background = element_rect(size = 1.6, fill = NA)
       , panel.border = element_rect(size = 1.6, fill=NA)
       , axis.text.x   = element_text( size   = gg_ptsz
       , margin = unit(rep(0.35,times=4),"cm"))#end element_text
       , axis.text.y       = element_text( size   = gg_ptsz
       , margin = unit(rep(0.35,times=4),"cm")
      )#end element_text
      , axis.ticks.length = unit(-0.25,"cm")
      ) #end theme
out_name = "summed-BF-qunt-obs.png"
ggsave(gg_now, filename=out_name,path=secy_path,width=gg_widthn*0.6,height=gg_height*0.6,dpi=gg_depth,device=gg_device[1])



```









```{r,label="plot-stand-structure-time-series"}

prcp_gradnt = mean_hlm %>% mutate(grdnt = case_when(rain > 100 ~ "wet",
                                                    rain > 50  & rain < 100 ~ "mid",
                                                    rain < 50 ~ "dry"))
prcp_gradnt = prcp_gradnt %>% mutate(sample = case_when (lat>40 & lat < 40.1 ~ "S"))
prcp_gradnt = prcp_gradnt %>% filter(!is.na(sample) & !is.na(grdnt))
prcp_gradnt = prcp_gradnt %>% group_by(grdnt) %>%  mutate(x = lon[which.max(elai)],
                                                          y = lat[which.max(elai)])
wet_x = round(prcp_gradnt$x[prcp_gradnt$grdnt=="wet"][1], 5)
wet_y = round(prcp_gradnt$y[prcp_gradnt$grdnt=="wet"][1], 5)
mid_x = round(prcp_gradnt$x[prcp_gradnt$grdnt=="mid"][1], 5)
mid_y = round(prcp_gradnt$y[prcp_gradnt$grdnt=="mid"][1], 5)
dry_x = round(prcp_gradnt$x[prcp_gradnt$grdnt=="dry"][1], 5)
dry_y = round(prcp_gradnt$y[prcp_gradnt$grdnt=="dry"][1], 5)


## slice stand structure for the three sites to plot out time series 
    
bydbh_wet = bydbh                                             %>% 
            mutate(lon=round(lon,5)
                 ,lat=round(lat,5))                           %>% 
           filter(lon == wet_x & lat == wet_y)                %>% 
           mutate(site = "Wet")                               %>% 
           mutate(fates_ddbh_canopy = fates_ddbh_canopy/fates_nplant_canopy,
                  fates_ddbh_ustory = fates_ddbh_ustory/fates_nplant_ustory)

bydbh_mid = bydbh                                             %>% 
            mutate(lon=round(lon,5)
                 ,lat=round(lat,5))                           %>% 
           filter(lon == mid_x & lat == mid_y)                %>% 
           mutate(site = "Mid")                               %>% 
           mutate(fates_ddbh_canopy = fates_ddbh_canopy/fates_nplant_canopy,
                  fates_ddbh_ustory = fates_ddbh_ustory/fates_nplant_ustory)

bydbh_dry = bydbh                                             %>% 
            mutate(lon=round(lon,5)
                 ,lat=round(lat,5))                           %>% 
           filter(lon == dry_x & lat == dry_y)                %>% 
            mutate(site = "Dry")                              %>% 
           mutate(fates_ddbh_canopy = fates_ddbh_canopy/fates_nplant_canopy,
                  fates_ddbh_ustory = fates_ddbh_ustory/fates_nplant_ustory)

site_bydbh = rbind(bydbh_wet, bydbh_mid, bydbh_dry)

#--- Title for legend
dbh_legend = desc.unit(desc="DBH",unit=untab$cm)
#---~---

dbh_loop = which(fatesvar_szpf$vnam %in% names(bydbh))
gg_dbh   = list()

  for (f in dbh_loop){
  #--- Match variables.
  f_vnam   = fatesvar_szpf$vnam [f]
  f_desc   = fatesvar_szpf$desc [f]
  f_unit   = fatesvar_szpf$unit [f]
  f_stack  = fatesvar_szpf$stack[f]
  #f_dbh01  = fatesvar_szpf$dbh01[f]
  f_dbh01  = FALSE #ignore the 0-5cm size class for oak
  cat0("   - ",f_desc,".")
  #---~---
  
  #--- Decide whether to plot the first class.
  if (f_dbh01){
    #--- Keep all classes.
    f_bydbh     = site_bydbh
    f_bydbh$dbh = factor(f_bydbh$dbh,levels=sequence(ndbhs))
    f_coldbhs   = magma(ndbhs,direction=1)
    f_dbhlabs   = dbhinfo$labs
    f_dbhlabs   = parse(text = f_dbhlabs)
    #---~---
  }else{
    #--- Exclude first class.
    bye         = as.numeric(site_bydbh$dbh) %in% 1
    f_bydbh     = site_bydbh[! bye,]
    f_bydbh$dbh = factor(f_bydbh$dbh,levels=sequence(ndbhs)[-1])
    f_coldbhs   = magma(ndbhs,direction=1)[-1]
    f_dbhlabs   = dbhinfo$labs[-1]
    f_dbhlabs   = parse(text = f_dbhlabs)
  }#end if (f_dbh01)
  #---~---
   
  #--- Initialise plot (decide whether to plot lines or stacks).
  if (f_stack){
     gg_now = ggplot(data=f_bydbh,aes_string(x="time",y=f_vnam,group="dbh",fill="dbh"))
     gg_now = gg_now + facet_wrap(.~site)
     gg_now = gg_now + scale_fill_manual(name=dbh_legend,labels=f_dbhlabs,values=f_coldbhs)
     gg_now = gg_now + geom_area(position=position_stack(reverse = FALSE),show.legend = TRUE)
  }else{
     gg_now = ggplot(data=f_bydbh,aes_string(x="time",y=f_vnam,group="dbh",colour="dbh"))
     gg_now = gg_now + scale_colour_manual(name=dbh_legend,labels=f_dbhlabs,values=f_coldbhs)
     gg_now = gg_now + geom_line(lwd=1.0,show.legend = TRUE)
  }#end if (f_stack)
  gg_now = gg_now + labs(title=case_desc)
  gg_now = gg_now + scale_x_datetime(date_labels=gg_tfmt)
  gg_now = gg_now + xlab("Simulation time")
  gg_now = gg_now + ylab(desc.unit(desc=f_desc,unit=untab[[f_unit]],twolines=TRUE))
  gg_now = gg_now + theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size = 0.5,base_rect_size =0.5)
  gg_now = gg_now + theme( legend.position   = "bottom"
                         , axis.text.x       = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.text.y       = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.ticks.length = unit(-0.25,"cm")
                         )#end theme
  #---~---

  #--- Save plot.
  for (d in sequence(ndevice)){
    f.output = paste0(f_vnam,"-tsdbh-rev",".",gg_device[d])
    dummy = ggsave( filename = f.output
                  , plot     = gg_now
                  , device   = gg_device[d]
                  , path     = tsdbh_path
                  , width    = gg_width
                  , height   = gg_height
                  , units    = gg_units
                  , dpi      = gg_depth
                  )#end ggsave
  }#end for (o in sequence(nout))
  #---~---

  #--- Write plot settings to the list.
  gg_dbh[[f_vnam]] = gg_now
  #---~---
}#end for (a in age_loop)



## histogram plots for NPLANT
site_nplant = site_bydbh %>% dplyr::select(site, time, dbh, fates_nplant)
site_nplant = site_nplant %>% group_by(site,dbh) %>% summarize(fates_nplant = mean(fates_nplant))
site_nplant = site_nplant %>%  arrange(site,dbh)
site_nplant$dbh = factor(site_nplant$dbh,levels=unique(site_nplant$dbh))
gg_now = ggplot(site_nplant, aes(dbh, fates_nplant)) 
gg_now = gg_now + facet_wrap(.~site) 
gg_now = gg_now + geom_bar(stat = "identity") 
gg_now = gg_now + labs(title=case_desc)
gg_now = gg_now + xlab("Size class index")
gg_now = gg_now + ylab(expression("( Plant"~Ha^-1~")"))
gg_now = gg_now + theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size = 0.5,base_rect_size =0.5)
gg_now = gg_now + theme( legend.position   = "bottom"
                         , axis.text.x       = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.text.y       = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.ticks.length = unit(-0.25,"cm")
                         )#

f.output = paste0("NPALNT","-bydbh-",".",gg_device[1])
dummy = ggsave(     filename = f.output
                  , plot     = gg_now
                  , device   = gg_device[d]
                  , path     = tsdbh_path
                  , width    = gg_width
                  , height   = gg_height
                  , units    = gg_units
                  , dpi      = gg_depth
                  )


```






```{r,label="plot-DDBH-annual-dynamic-by-site"}


site_ddbh  = site_bydbh                        %>%    
             dplyr::select(site,
                           time,
                           dbh,
                           fates_ddbh_canopy,
                           fates_ddbh_ustory)  %>% 
             mutate(year = year(time))         %>% 
             group_by(site,year,dbh)           %>% 
             summarize(fates_ddbh_canopy = mean(fates_ddbh_canopy),
                       fates_ddbh_ustory = mean(fates_ddbh_ustory)) %>% 
             ungroup()                         %>% 
             mutate(year = as.POSIXct(ymd(year, truncated = 2L)))

dbh_loop = which(fatesvar_szpf$vnam %in% names(site_ddbh))
gg_dbh   = list()

  for (f in dbh_loop){
  #--- Match variables.
  f_vnam   = fatesvar_szpf$vnam [f]
  f_desc   = fatesvar_szpf$desc [f]
  #f_unit   = fatesvar_szpf$unit [f]
  f_unit   = "cmoployr"
  f_stack  = fatesvar_szpf$stack[f]
  f_stack  = FALSE
  #f_dbh01  = fatesvar_szpf$dbh01[f]
  f_dbh01  = FALSE #ignore the 0-5cm size class for oak
  cat0("   - ",f_desc,".")
  #---~---
  
  #--- Decide whether to plot the first class.
  if (f_dbh01){
    #--- Keep all classes.
    f_bydbh     = site_ddbh
    f_bydbh$dbh = factor(f_bydbh$dbh,levels=sequence(ndbhs))
    f_coldbhs   = magma(ndbhs,direction=1)
    f_dbhlabs   = dbhinfo$labs
    f_dbhlabs   = parse(text = f_dbhlabs)
    #---~---
  }else{
    #--- Exclude first class.
    bye         = as.numeric(site_ddbh$dbh) %in% 1
    f_bydbh     = site_ddbh[! bye,]
    f_bydbh$dbh = factor(f_bydbh$dbh,levels=sequence(ndbhs)[-1])
    f_coldbhs   = magma(ndbhs,direction=1)[-1]
    f_dbhlabs   = dbhinfo$labs[-1]
    f_dbhlabs   = parse(text = f_dbhlabs)
  }#end if (f_dbh01)
  #---~---
   
  #--- Initialise plot (decide whether to plot lines or stacks).
  if (f_stack){
     gg_now = ggplot(data=f_bydbh,aes_string(x="year",y=f_vnam,group="dbh",fill="dbh"))
     gg_now = gg_now + facet_wrap(.~site, nrow=3, scales="free")
     gg_now = gg_now + scale_fill_manual(name=dbh_legend,labels=f_dbhlabs,values=f_coldbhs)
     gg_now = gg_now + geom_area(position=position_stack(reverse = FALSE),show.legend = TRUE)
  }else{
     gg_now = ggplot(data=f_bydbh,aes_string(x="year",y=f_vnam,group="dbh",colour="dbh"))
     gg_now = gg_now + facet_wrap(.~site, nrow=3, scales="free")
     gg_now = gg_now + scale_colour_manual(name=dbh_legend,labels=f_dbhlabs,values=f_coldbhs)
     gg_now = gg_now + geom_line(lwd=1.0,show.legend = TRUE)
  }#end if (f_stack)
  gg_now = gg_now + labs(title=case_desc)
  gg_now = gg_now + scale_x_datetime(date_labels=gg_tfmt)
  gg_now = gg_now + xlab("Simulation time")
  #gg_now = gg_now + ylab(desc.unit(desc=f_desc,unit=untab[[f_unit]],twolines=TRUE))
  gg_now = gg_now + ylab(bquote(.(f_desc)~"("~cm~plant^-1~yr^-1~")"))
  gg_now = gg_now + theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size = 0.5,base_rect_size =0.5)
  gg_now = gg_now + theme( legend.position   = "bottom"
                         , axis.text.x       = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.text.y       = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.ticks.length = unit(-0.25,"cm")
                         )#end theme
  #---~---

  #--- Save plot.
  for (d in sequence(ndevice)){
    f.output = paste0(f_vnam,"-yeardbh-rev",".",gg_device[d])
    dummy = ggsave( filename = f.output
                  , plot     = gg_now
                  , device   = gg_device[d]
                  , path     = tsdbh_path
                  , width    = gg_width
                  , height   = gg_height
                  , units    = gg_units
                  , dpi      = gg_depth
                  )#end ggsave
  }#end for (o in sequence(nout))
  #---~---

  #--- Write plot settings to the list.
  gg_dbh[[f_vnam]] = gg_now
  #---~---
}#end for (a in age_loop)

```




```{r,label="correlation-matrix"}

cor_all  = mean_allseson
cor_peak = mean_pseason


type   = "peak season"
cor_peak = cor_peak[,c(1,4:5,7,9:10,12:15,18:19)]
nam_peak = c("GPP","AGB","LAI","$ Rain[grow]","$ AirTemp[2*m-grow]",
              "$ qv[2*m-grow]", "BurnFrac", "FuelLoad","$ qv[2*m-dry]",
               "$ AirTemp[2*m-dry]","$ WindSpeed[10*m]")


for(c in sequence(ncase)){
  case_now  = case_names[c]
  desc_now  = case_desc[c]
  cor_now   = cor_peak                    %>% 
              filter(case == case_now)  %>% 
              dplyr::select(-case)      %>% 
              filter(!is.na(fates_gpp))

  corr_allens       = cor(cor_now)
  p_all             = cor.mtest(cor_now,conf.level=0.95, method="pearson")$p
  
  colnames(corr_allens) = nam_peak
  rownames(corr_allens) = nam_peak

  colnames(p_all) = nam_peak
  rownames(p_all) = nam_peak
  
  f_output = paste0(desc_now,"_",type,"_correlation.png")
  png(file=file.path(plot_main,f_output),width=13,height=8,units="in",res=600)

  corrplot(corr_allens
          ,method      = "ellipse"
          ,type        = "lower"
          ,p.mat       = p_all
          ,sig.level   = 0.05
          ,insig       = "blank"
          ,addCoef.col = "grey50"
          ,title = paste(desc_now,type,"correlation",sep=" ")
          ,mar=c(0,0,1,0)
          ,order = "original"
          ,diag  = FALSE
          ,number.cex=0.55
          ,tl.cex = 1
          ,tl.col="black"
          ,cl.cex=1)
  dev.off()
}


cor_all   = cor_all[,c(1,4:5,7,9:10,12:14,16)]
nam_all   = c("GPP","AGB","LAI","$ Rain", "$ AirTemp[2*m]","$ qv[2*m]",
              "BurnFrac", "FuelLoad","$ WindSpeed[10*m]")

type   = "all season"

for(c in sequence(ncase)){
  case_now  = case_names[c]
  desc_now  = case_desc[c]
  cor_now   = cor_all                    %>% 
              filter(case == case_now)   %>% 
              dplyr::select(-case)       %>% 
              filter(!is.na(fates_gpp))

  corr_allens       = cor(cor_now)
  p_all             = cor.mtest(cor_now,conf.level=0.95, method="pearson")$p
  
  colnames(corr_allens) = nam_all
  rownames(corr_allens) = nam_all

  colnames(p_all) = nam_all
  rownames(p_all) = nam_all
  
  f_output = paste0(desc_now,"_",type,"_correlation.png")
  png(file=file.path(plot_main,f_output),width=6,height=5,units="in",res=600)

  corrplot(corr_allens
          ,method      = "ellipse"
          ,type        = "lower"
          ,p.mat       = p_all
          ,sig.level   = 0.05
          ,insig       = "blank"
          ,addCoef.col = "grey50"
          #,title = paste(desc_now,type,"correlation",sep=" ")
          ,mar=c(0,0,1,0)
          ,order = "original"
          ,diag  = FALSE
          ,number.cex=0.55
          ,tl.cex = 1
          ,tl.col="black"
          ,cl.cex=1)
  dev.off()
}



```








```{r,label="variance-decomposition-main"}

if_plot = FALSE
sens_y    = c("fates_gpp","elai","fates_burnfrac")
sens_xp   = c("fates_fuel_amount","rain","tsa_g","q2m_g",
              "tsa_d","q2m_d","u10")
sens_xa   =c("fates_fuel_amount","rain","tsa","q2m","u10") 


sens_df   = mean_allseson
type      = "AllSeason"
sens_x    = sens_xa

nx        = length(sens_x)
ny        = length(sens_y)
sens_main = matrix( data=NA_real_
                  , nrow=ny*ncase
                  , ncol=nx+1
                  , dimnames = list(rep(sens_y,time=ncase),c(sens_x,"case")))
last_col = ncol(sens_main)
sens_plot = list()

for(c in sequence(ncase)){
  case_now  = case_names[c]
  desc_now  = case_desc[c]
  sens_now  = sens_df   %>%  filter(case==case_now) %>% filter(!is.na(fates_burnfrac))
  for(v in sequence(ny)){
  var_now = sens_y[v]
  for(p in sequence(nx)){
    p_now   = sens_x[p] 
    df_now  = sens_now %>% select(p_now,var_now)
    x       = df_now[[p_now]]
    y       = df_now[[var_now]]
    fit     = lm(y ~ x)
    moddf   = data.frame(x=x,y=y)
    moddf$pred = predict(fit,data.frame(x=x))
    if(if_plot){
      
      gg_now = ggplot(data=moddf,aes(x,y))
      gg_now = gg_now + geom_point()
      gg_now = gg_now + geom_line(data=moddf,aes(x,pred),lwd=1)
      gg_now = gg_now + xlab(rlang::sym(p_now))
      gg_now = gg_now + ylab(rlang::sym(var_now))
      gg_now = gg_now + theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size = 0.5,base_rect_size =0.5)
      gg_now = gg_now + theme( legend.position   = "bottom"
                             , axis.text.x       = element_text( size   = gg_ptsz
                                                               , margin = unit(rep(0.35,times=4),"cm")
                                                               )#end element_text
                             , axis.text.y       = element_text( size   = gg_ptsz
                                                               , margin = unit(rep(0.35,times=4),"cm")
                                                               )#end element_text
                             , plot.title        = element_text( size = gg_ptsz)
                             , axis.ticks.length = unit(-0.25,"cm")
                             )#end theme
    
     for (d in sequence(ndevice)){
     f_output = paste0(desc_now,"-",type,"-",var_now,"-",p_now,".",gg_device[d])
     dummy    = ggsave( filename = f_output
                     , plot     = gg_now
                     , device   = gg_device[d]
                     , path     = sens_path
                     , width    = gg_width
                     , height   = gg_height
                     , units    = gg_units
                     , dpi      = gg_depth
                     )
          } #end of for(d in sequence(ndevice))
  sens_plot[[desc_now]][[var_now]][[p_now]]=gg_now
    } # end of if(if_plot)
  if(c==1){
  sens_main[v,p]           = var(moddf$pred)/var(moddf$y)
  sens_main[v,last_col]    = desc_now  
  }else{
  v_idx = (c-1)*ny   + v
  sens_main[v_idx,p]           = var(moddf$pred)/var(moddf$y)
  sens_main[v_idx,last_col]    = desc_now 
       }#end if else
 
      } #end p in sequence(nx)   
   } #end v in sequence(ny)
} # end c in sequnece(ncase)



```








```{r,label="variance-decomposition-min"}

sens_min = matrix( data=NA_real_
                  ,nrow=ny*ncase
                  ,ncol=nx+1
                  ,dimnames = list(rep(sens_y,time=ncase),c(sens_x,"case")))
sens_plot = list()

for(c in sequence(ncase)){
  case_now  = case_names[c]
  desc_now  = case_desc[c]
  sens_now  = sens_df   %>%  filter(case==case_now) %>% filter(!is.na(fates_burnfrac))
  for(v in sequence(ny)){
  var_now   = sens_y[v]
  y         = sens_now[[var_now]]
  for(p in sequence(nx)){
    p_now   = sens_x[p] 
    x       = sens_now[[p_now]]
    y_rsd   = y
    others  = sens_x[sens_x!=p_now]
    n_other = length(others)
    for(t in sequence(n_other)){
      other_var = others[t]
      x_other   = sens_now[[other_var]]
      bs_x      = x_other
      bs_y      = y_rsd
      fit       = lm(bs_y ~ bs_x)
      pred      = predict(fit,data.frame(bs_x=bs_x))
      y_rsd     = y_rsd - pred
    }#end t in sequence(n_other)
    fit_x     = lm(y_rsd ~ x)
    x_pred    = predict(fit_x,data.frame(x = x))
    plotdf    = data.frame(x=x,y=y_rsd,pred=x_pred,orig_y = y)
    var_min   = var(plotdf$pred)/var(plotdf$orig_y)
    
    if(c==1){
     sens_min[v,p] = var_min 
     sens_min[v,last_col]=desc_now
    }else{
     v_idx = (c-1)*ny + v
     sens_min[v_idx,p]  = var_min
     sens_min[v_idx,last_col] = desc_now
    } #end if else
    
   if(if_plot){
    gg_now = ggplot(data=plotdf,aes(x,y))
    gg_now = gg_now + geom_point()
    gg_now = gg_now + geom_line(data=plotdf,aes(x,pred),lwd=1)
    gg_now = gg_now + xlab(rlang::sym(p_now))
    gg_now = gg_now + ylab(paste0(var_now,"_residual"))
    gg_now = gg_now + theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size = 0.5,base_rect_size =0.5)
    gg_now = gg_now + theme( legend.position   = "bottom"
                         , axis.text.x       = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.text.y       = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , plot.title        = element_text( size = gg_ptsz)
                         , axis.ticks.length = unit(-0.25,"cm")
                         )#end theme
    
    for (d in sequence(ndevice)){
    f_output = paste0(desc_now,"-",type,"-",var_now,"-",p_now,"-varmin.",gg_device[d])
    dummy = ggsave( filename = f_output
                  , plot     = gg_now
                  , device   = gg_device[d]
                  , path     = sens_path
                  , width    = gg_width
                  , height   = gg_height
                  , units    = gg_units
                  , dpi      = gg_depth
                  )
     } #end d in sequence(ndevice)
  sens_plot[[desc_now]][[var_now]][[p_now]]=gg_now 
    } #end of if(if_plot)
   } #end p in sequence(nx)
  } #end v in sequence(ny)
} #end c in sequence(ncase)



```







```{r,label="effects-ranking",fig.height=9,fig.width=11.25}

var                  = rownames(sens_main)
sens_madf            = as_tibble(sens_main)
sens_madf$model_var  = var
sens_madf$type       = "main"

var                  = rownames(sens_min)
sens_midf            = as_tibble(sens_min)
sens_midf$model_var  = var 
sens_midf$type       = "min"
sens_decom           = rbind(sens_madf,sens_midf)
sens_decom           = sens_decom                                           %>%  
                       pivot_longer(cols = fates_fuel_amount:u10
                                   ,names_to  = "parameter"
                                   ,values_to = "var_decomp")   

#par_label            = nam_peak[c(8,4:6,10,9,11)]
par_label            = nam_all[c(8,4:6,9)]
par_label            = gsub(pattern="\\$ ", replacement="",x=par_label)

sens_decom           = sens_decom                                                %>% 
                       mutate( var_decomp = round(as.numeric(var_decomp)*100,1)
                             , label      = rep(par_label,time=ny*ncase*2)) 
  
var_label            = c("GPP"
                        ,"LAI"
                        ,"Burned fraction"
                        )

names(var_label)     = c("fates_gpp"
                        ,"elai"
                        ,"fates_burnfrac"
                        )


decomp_plot=list()
for(c in sequence(ncase)){
  desc_now   = case_desc[c]
  decomp_now = sens_decom %>% filter(case==desc_now & model_var=="fates_burnfrac") 
  
  gg_now = ggplot(decomp_now,aes(var_decomp,label)) 
  gg_now = gg_now + geom_linerange(aes(xmin=0,xmax=var_decomp,colour=type)
                                  ,size=4, position = position_dodge(width = 0.5))
  gg_now = gg_now + scale_colour_manual(values = c("black","grey50"))
  gg_now = gg_now + facet_wrap(.~ model_var,labeller=labeller(model_var=var_label))
  #gg_now = gg_now + labs(title = paste(desc_now,type,sep="-"))
  gg_now = gg_now + xlab("Variance explained (%)")
  gg_now = gg_now + ylab(element_blank())
  gg_now = gg_now + scale_y_discrete(labels = scales::label_parse(),limits=rev)
  gg_now = gg_now + theme_grey( base_size = gg_ptsz
                              , base_family = "Helvetica"
                              , base_line_size = 0.5
                              , base_rect_size = 0.5)
  gg_now = gg_now +  theme( axis.text.x = element_text( size   = gg_ptsz
                                                      , margin = unit(rep(0.35,times=4),"cm")
                                                      )#end element_text
                          , axis.text.y = element_text( size   = gg_ptsz-5
                                                      , margin = unit(rep(0.35,times=4),"cm")
                                                      )#end element_text
                          , axis.ticks.length = unit(-0.25,"cm")
                          , legend.position   = "bottom"
                          , legend.direction  = "horizontal"
                          , plot.margin = unit(c(0,0,0,0),"cm")
                          )#end theme
  for (d in sequence(ndevice)){
    f_output = paste0(desc_now,"-",type,"-variance-decomp.",gg_device[d])
    dummy    = ggsave( filename = f_output
                     , plot     = gg_now
                     , device   = gg_device[d]
                     , path     = plot_main
                     , width    = gg_widthn*0.4
                     , height   = gg_height*0.4
                     , units    = gg_units
                     , dpi      = gg_depth
                     )
     } #end d in sequence(ndevice)
 decomp_plot[[desc_now]] = gg_now
}
if(gg_screen) decomp_plot




```





```{r,label="multi-variates-model"}

lmod_df   = mean_allseson %>% filter(!is.na(fates_burnfrac)) %>% mutate(case = as.factor(case))
fuel_mod  = lm(fates_burnfrac ~ case*fates_fuel_amount, data = lmod_df)
summary(fuel_mod)

resid_fuel = resid(fuel_mod)

lmod_df$bf_rsd = resid_fuel
u10_mod   = lm(bf_rsd ~ u10, data = lmod_df)
rain_mod  = lm(bf_rsd ~ rain, data = lmod_df)
tsa_mod   = lm(bf_rsd ~ tsa, data = lmod_df)
q2_mod    = lm(bf_rsd ~ q2m, data = lmod_df)
full_mod  = lm(bf_rsd ~ rain*tsa*q2m*u10, data = lmod_df)
t = BIC(u10_mod,rain_mod,tsa_mod,q2_mod,full_mod)
summary(bf_mod)




```





```{r,label="AGB-climate-relationship-benchmark"}

## load pepper wood preserve data
agb_p = fread(agbP_path)
agbp_amean = agb_p                                                    %>% 
        mutate( site       = "pepwd"
              , latitude   = 38.56994
              , longitude  = -122.68958
              , date       = mdy(CollectDate)
              , year       = year(date)
              , biomass    = (FreshDryWt - FreshBagWt)*10*g2kg/2
              , litter     = (ThatchDryWt - ThatchBagWt)*10*g2kg/2)   %>% 
         dplyr::select(site,longitude,latitude,year,biomass,litter)   %>% 
         group_by(site,longitude,latitude,year)                       %>% 
         summarise(agb = mean(biomass,na.rm=TRUE)
                  ,agb_qupr = quantile(biomass,probs=qupr_ribbon
                                        ,na.rm=TRUE)
                  ,agb_qlwr = quantile(biomass,probs=qlwr_ribbon
                                        ,na.rm=TRUE))                 %>% 
         ungroup()                                                     
         

## load seabloom et al. data
## California sites: elliot.us, hopl.us, mcla.us,sage.us,sedg.us, and sier.us
ca_sites = c("elliot.us","hopl.us","mcla.us","sage.us","sedg.us","sier.us")
agb_s = fread(agbS_path,)
agbs_amean =   agb_s                                          %>% 
               filter(site %in% ca_sites & trt == "Control")  %>% 
               mutate(live = (10^live_lg)*g2kg/2)             %>% 
               group_by(site,longitude,latitude,year)         %>% 
               summarize(agb      = mean(live,na.rm=TRUE)
                        ,agb_qupr = quantile(live,probs=qupr_ribbon
                                             ,na.rm=TRUE)
                        ,agb_qlwr = quantile(live,probs=qlwr_ribbon
                                             ,na.rm=TRUE))    %>% 
               ungroup()

agb_amean = rbind(agbp_amean,agbs_amean)

## extract historical climate (annual temperature and precipitation) from worldclim for each site
mat   = raster(file.path(clim_main, "wc2.1_10m_bio_1.tif")) #mean annual temp
map   = raster(file.path(clim_main, "wc2.1_10m_bio_12.tif"))#mean annual precip 
mat_wetq = raster(file.path(clim_main, "wc2.1_10m_bio_8.tif")) # mean temp of wettest quarter
map_wetq = raster(file.path(clim_main, "wc2.1_10m_bio_16.tif")) #precip of wettest quarter
ca_xy = raster::extent(-124.5,-114.5, 32,42.5)
ca_rs = raster(ca_xy,res=0.1666667,crs=wgs)
vals  = 1:ncell(ca_rs)
ca_rs = setValues(ca_rs,vals)
mat   = crop(mat,ca_rs)
map   = crop(map,ca_rs)
mat_wetq = crop(mat_wetq,ca_rs)
map_wetq = crop(map_wetq,ca_rs)
mat_df = as.data.frame(mat,xy=TRUE) %>% rename(mat = wc2.1_10m_bio_1)
map_df = as.data.frame(map,xy=TRUE) %>% rename(map = wc2.1_10m_bio_12)
matq_df = as.data.frame(mat_wetq,xy=TRUE) %>% rename(mat_wetq =wc2.1_10m_bio_8)
mapq_df = as.data.frame(map_wetq,xy=TRUE) %>% rename(map_wetq =wc2.1_10m_bio_16)

## search for the nearest grid to extrac climate information
nn_pt = RANN::nn2(mat_df[,1:2],agb_amean[,2:3],k=1)
agb_amean$id =  as.vector(nn_pt$nn.idx)
agb_amean    = agb_amean                         %>% 
               mutate(mat = mat_df$mat[id]
                     ,map = map_df$map[id]
                     ,mat_wetq = matq_df$mat_wetq[id]
                     ,map_wetq = mapq_df$map_wetq[id])      %>% 
               dplyr::select(-id)

mod_agb     = hlm1d                         %>% 
              filter(case==2)               %>% 
              mutate(year=year(time))       %>% 
              group_by(lon_f,lat_f,year)    %>% 
              summarize(agb_mod = max(fates_vegc_aboveground)
                       ,mat_mod = mean(tsa)
                       ,map_mod = mean(rain)*12)%>% 
              ungroup()
mod_xy      = hlm1d                                %>% 
              dplyr::select(lon,lat,lon_f,lat_f)   %>% 
              distinct()
nn_pt = RANN::nn2(mod_xy[,1:2],agb_amean[,2:3],k=1)
mod_site = agb_amean %>% dplyr::select(longitude,latitude)
mod_site$id = as.vector(nn_pt$nn.idx)   
mod_site    =  mod_site               %>% 
               mutate(lon = mod_xy$lon[id]
                     ,lat = mod_xy$lat[id]
                     ,flag = TRUE
                     )

mod_df  = mod_xy %>% left_join(mod_site,by=c("lon","lat")) %>% filter(flag) 

mod_df  = mod_agb %>% left_join(mod_df,by=c("lon_f","lat_f")) %>% filter(flag) %>% distinct() 


site_lab = agb_amean %>% dplyr::select(site,longitude,latitude) %>% distinct()
mod_df   = mod_df %>% left_join(site_lab,by=c("longitude","latitude")) %>% 
           dplyr::select(year,site,agb_mod,mat_mod,map_mod) %>% 
           mutate(source="model") %>% rename(agb=agb_mod,mat=mat_mod,map=map_mod)

obs_df   = agb_amean %>% dplyr::select(year,site,agb,mat,map) %>% mutate(source="site")
plot_df = rbind(obs_df,mod_df)           


gg_now = ggplot(data=plot_df,aes(x=mat,y=agb,colour=source)) + 
         geom_point()  +
         geom_smooth(method="lm",size=1) +
         labs( x=expression("Mean annual temperature ("*degree~C*")")
             , y="Peak live aboveground biomass (kgC)")+
         theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size= 0.5,base_rect_size =0.5)+
         theme(  legend.position   = "right" 
               , legend.title=element_blank()
               , panel.background  = element_blank()
               , panel.border      = element_rect(linewidth = 1.6, fill=NA)
               , axis.text.x       = element_text( size   = gg_ptsz-5
                                                 , margin = unit(rep(0.35,times=4),"cm"))#end element_text
               , axis.text.y       = element_text( size   = gg_ptsz-5
                                                 , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
               , axis.ticks.length = unit(-0.25,"cm")
                ) #end theme
ggsave(gg_now,filename="agb-mat-benchmark.png",path=plot_main,
       width=gg_widthn*0.6,height=gg_height*0.6,unit=gg_units,dpi=gg_depth)         


gg_now = ggplot(data=plot_df,aes(x=map,y=agb,colour=source)) + 
         geom_point()  +
         geom_smooth(method="lm",size=1) +
         labs( x="Mean annual precipitation (mm)"
             , y="Peak live aboveground biomass (kgC)")+
         theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size= 0.5,base_rect_size =0.5)+
         theme(  legend.position   = "right" 
               , legend.title=element_blank()
               , panel.background  = element_blank()
               , panel.border      = element_rect(linewidth = 1.6, fill=NA)
               , axis.text.x       = element_text( size   = gg_ptsz-5
                                                 , margin = unit(rep(0.35,times=4),"cm"))#end element_text
               , axis.text.y       = element_text( size   = gg_ptsz-5
                                                 , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
               , axis.ticks.length = unit(-0.25,"cm")
                ) #end theme

ggsave(gg_now,filename="agb-map-benchmark.png",path=plot_main,
       width=gg_widthn*0.6,height=gg_height*0.6,unit=gg_units,dpi=gg_depth)


gg_now = ggplot(data=plot_df,aes(x=mat,y=map,colour=source)) + 
         geom_point()  +
         geom_smooth(method="lm",size=1) +
         labs( x=expression("Mean annual temperature ("*degree~C*")")
             , y="Mean annual precipitation (mm)")+
         theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size= 0.5,base_rect_size =0.5)+
         theme(  legend.position   = "right" 
               , legend.title=element_blank()
               , panel.background  = element_blank()
               , panel.border      = element_rect(linewidth = 1.6, fill=NA)
               , axis.text.x       = element_text( size   = gg_ptsz-5
                                                 , margin = unit(rep(0.35,times=4),"cm"))#end element_text
               , axis.text.y       = element_text( size   = gg_ptsz-5
                                                 , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
               , axis.ticks.length = unit(-0.25,"cm")
                ) #end theme
ggsave(gg_now,filename="mat-map-correlation.png",path=plot_main,
       width=gg_widthn*0.6,height=gg_height*0.6,unit=gg_units,dpi=gg_depth) 


gg_now = ggplot(plot_df,aes(year,agb,colour=source)) + geom_line() +
         facet_wrap(.~site) +
         labs( x=""
             , y="Peak live aboveground biomass (kgC)")+
         theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size= 0.5,base_rect_size =0.5)+
         theme(  legend.position   = "right" 
               , legend.title=element_blank()
               , panel.background  = element_blank()
               , panel.border      = element_rect(linewidth = 1.6, fill=NA)
               , axis.text.x       = element_text( size   = gg_ptsz-5
                                                 , margin = unit(rep(0.35,times=4),"cm"))#end element_text
               , axis.text.y       = element_text( size   = gg_ptsz-5
                                                 , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
               , axis.ticks.length = unit(-0.25,"cm")
                ) #end theme
         
ggsave(gg_now,filename="agb-by-year-benchmark.png",path=plot_main,
       width=gg_widthn*0.6,height=gg_height*0.6,unit=gg_units,dpi=gg_depth) 



```








```{r,label="load-site-data"}
if ("nc_site" %in% ls()){dummy = nc_close(nc_site); rm(nc_site)}

# TO-DO: loop for reading multiple site data
# Open NetCDF connection and retrieve variable names
cat0(" + Load site data from ",site_base,".")
nc_site  = nc_open(filename=site_file)
nc_nvars = nc_site$nvars
nc_ndims = nc_site$ndims
nc_dlist = rep(NA_character_,times=nc_ndims)
nc_vlist = rep(NA_character_,times=nc_nvars)
for (d in sequence(nc_ndims)) nc_dlist[d] = nc_site$dim[[d]]$name
for (v in sequence(nc_nvars)) nc_vlist[v] = nc_site$var[[v]]$name

# Select variables to load
nc_obs1d    = nc_vlist[tolower(nc_vlist) %in% hlm1dvar$vnam[hlm1dvar$assess]]
nc_obs2d    = nc_vlist[tolower(nc_vlist) %in% hlm2dsoi$vnam[hlm2dsoi$assess]]
nc_obs      = c(nc_obs1d, nc_obs2d)


# Extract time information
site_time0  = as_datetime(gsub(pattern="^days since ",replacement="",x=nc_site$dim$time$units))
site_time   = site_time0 + days(nc_site$dim$time$vals)
n_site_time = nc_site$dim$time$len

# Initialise a tibble that will host all data
site1d = tibble( time = site_time)

# Find conversion factors for monthly variables.
cmon.day = days_in_month(site1d$time)
cmon.hr  = day.hr  * cmon.day
cmon.min = day.min * cmon.day
cmon.sec = day.sec * cmon.day

# Loop through variables, and load data sets.
for (o in seq_along(nc_obs)){
   nc_nvnow        = nc_obs[o]
   if(nc_nvnow=="H2OSOI"){ #in hlm2dsoi
   h               = match(tolower(nc_nvnow),hlm2dsoi$vnam)
   h_vnam          = hlm2dsoi$vnam[h]
   h_desc          = hlm2dsoi$desc[h]
   h_add0          = eval(parse(text=hlm2dsoi$add0_sl[h]))
   h_mult          = eval(parse(text=hlm2dsoi$mult_sl[h]))
   nc_dat          = ncvar_get(nc=nc_site,varid=nc_nvnow)
   cat0("   - Retrieve ",h_desc,".")
   site1d[[h_vnam]] = h_add0 + h_mult * nc_dat  
   }else{
   h               = match(tolower(nc_nvnow),hlm1dvar$vnam)
   h_vnam          = hlm1dvar$vnam[h]
   h_desc          = hlm1dvar$desc[h]
   h_add0          = eval(parse(text=hlm1dvar$add0[h]))
   h_mult          = eval(parse(text=hlm1dvar$mult[h]))
   nc_dat          = ncvar_get(nc=nc_site,varid=nc_nvnow)
   cat0("   - Retrieve ",h_desc,".")
   site1d[[h_vnam]] = h_add0 + h_mult * nc_dat
   }
}#end for (o in seq_along(nc_obs))


# Load site coordinates. We will check whether observations and model are reasonably close.
site_coord = tibble( clon = c(ncvar_get(nc=nc_site,varid="LONGXY"))
                   , clat = c(ncvar_get(nc=nc_site,varid="LATIXY"))
                   , wlon = c(ncvar_get(nc=nc_site,varid="EDGEW" ))
                   , elon = c(ncvar_get(nc=nc_site,varid="EDGEE" ))
                   , slat = c(ncvar_get(nc=nc_site,varid="EDGES" ))
                   , nlat = c(ncvar_get(nc=nc_site,varid="EDGEN" ))
                   )#end tibble


# Close file and remove connection.
dummy = nc_close(nc_site)
rm(nc_site)

```






```{r,label="time-intersect"}
# Find the minimum interval that has overlaps with both site and model.
tstamp_each = unique(ens_tstamp)

tstampa_com = max(c(min(tstamp_each),min(site1d$time)))
tstampz_com = min(c(max(tstamp_each),max(site1d$time)))

# Restrict time to the overlapping period
tstamp_com  = tstamp_each[(tstamp_each >= tstampa_com) & (tstamp_each <= tstampz_com)]
ntstamp_com = length(tstamp_com)

# Trim site observations to the overlapping period. Also, standardize the data so we
# use only one of NA or NaN.
site = site1d %>%
         filter( (time >= tstampa_com) & (time <= tstampz_com)) %>%
         mutate_at(vars(-time),function(x) ifelse(is.nan(x),NA,x))

#  Load site LAI and AGB to bind with site
lai = fread(site_lai)
lai = lai                                %>% 
      mutate(LAI_DATE = ymd(LAI_DATE)
            ,year     = year(LAI_DATE)
            ,month    = month(LAI_DATE)) %>% 
      dplyr::select(-LAI_DATE)           %>% 
      group_by(year,month)               %>% 
      summarize_all(mean,na.rm=TRUE)     %>% 
      filter(!is.na(year))               

colnames(lai) <- c("year","month","elai","elai_sd")
agb = fread(site_agb)
agb = agb                                 %>% 
      mutate(date       = ymd(date)
            ,year       = year(date)
            ,month      = month(date)
            ,biomass    = biomass/2
            ,biomass_sd = biomass_sd/2)   %>% 
      dplyr::select(-date)                %>% 
      group_by(year,month)                %>% 
      summarise_all(mean,na.rm=TRUE)      %>% 
      ungroup()
colnames(agb) <- c("year","month","agb","agb_sd","sample")

## combine site observations
site  = site                                  %>%  
        mutate(year  = year(time)
              ,month = month(time))           %>% 
        left_join(agb,by=c("year","month"))   %>% 
        left_join(lai,by=c("year","month"))   %>% 
        dplyr::select(-c(agb_sd,sample,elai_sd))

site_var = names(site)
site_var = gsub(pattern="gpp",replacement="fates_gpp",x=site_var)
site_var = gsub(pattern="agb",replacement="fates_vegc_aboveground",x=site_var)
names(site) = site_var


```







```{r,label="merge-data",message=FALSE,results='hide'}
# First we have to filter out HLM variables that are not available at site or within the time frame of obs; 
# and then combine HLM with site obs to make comparison plots

cat0(" + Find the nearest grid cell from model simulation.")

site_xy = site_coord [1:2] %>% mutate(clon=clon-360)
nn_pt   = RANN::nn2(orig_xy[,2:3],site_xy[,1:2],k=1)
xy_model= orig_xy[nn_pt$nn.idx]
xy_model = xy_model %>% mutate(lon=round(lon,5),lat=round(lat,5))

    
hlm_site = hlm1d                                              %>% 
           mutate(lon=round(lon,5)
                 ,lat=round(lat,5))                           %>% 
           filter(lon == xy_model$lon & lat == xy_model$lat)   
   
          

# intersect variables
cat0(" + Find the variables common to both site and model.")
# Find common variables
var_both = intersect(names(site),names(hlm_site))
var_both = var_both[! var_both %in% "time"]
var_both = var_both[c(3,4,6,8,9)] # only select 5

# Restrict variables for both site and model
site  = site %>% dplyr::select( all_of(c("time",var_both)))

hlm_site = hlm_site                                                      %>% 
           dplyr::select( all_of(c('time','case',var_both)))  

cat0(" + Merge data sets into a single tibble.")

emean_com  = rbind( site %>% mutate(case = 0L)
                  , hlm_site)   %>%
             rename(source = case)


# Find the mean seasonal cycle.
mmean_com = emean_com                                                     %>%
            mutate(month = month(time))                                   %>%
            group_by(month,source)                                        %>%
            dplyr::select(-time)                                          %>%
            summarise_all(mean, na.rm=TRUE)                               %>%
            ungroup()                                                     %>%
            rename_at(vars(var_both), function(x) paste0(x,"_mean"))

# Find the lower range of the seasonal cycle
mqlwr_com = emean_com                                                         %>%
            mutate(month = month(time))                                       %>%
            group_by(month,source)                                            %>%
            dplyr::select(-time)                                              %>%
            summarise_all(quantile,probs=qlwr_ribbon,names=FALSE, na.rm=TRUE) %>%
            ungroup()                                                         %>%
            rename_at(vars(var_both), function(x) paste0(x,"_qlwr"))

# Find the upper range of the seasonal cycle
mqupr_com = emean_com                                                         %>%
            mutate(month = month(time))                                       %>%
            group_by(month,source)                                            %>%
            dplyr::select(-time)                                              %>%
            summarise_all(quantile,probs=qupr_ribbon,names=FALSE, na.rm=TRUE) %>%
            ungroup()                                                         %>%
            rename_at(vars(var_both), function(x) paste0(x,"_qupr"))

# Find the annual averages.
mmean_com = as_tibble( merge( x  = merge(x=mmean_com,y=mqlwr_com,by=c("month","source"))
                        , y  = mqupr_com
                        , by = c("month","source")
                        )#end merge
                 )#end as_tibble

# Sort data by month and source
mmean_com = mmean_com %>% arrange(source,month)         


```








```{r,label="model-site-comparison",message=FALSE,results='hide',fig.width=15, fig.height=12}
cat0(" + Plot the mean seasonal cycle of site measurements/estimates and model predictions.")

# Set legends
leg_colours = c(site_colour,ens_colour)
#leg_labels  = c("Site",as.character(case_desc))

gg_emean = list()
for( n in sequence(ncase)){
  case_now = case_names[n]
  desc_now = case_desc[[n]]
  h_mmean        = mmean_com %>% filter(source%in%c(0,case_now))
  leg_labels      = c("Site",desc_now)
  for (v in seq_along(var_both)){
   # Load variable
   h        = c(match(var_both[v],hlm1dvar$vnam), match(var_both[v],hlm2dsoi$vnam))
   hvars = c("hlm1dvar", "hlm2dsoi")
   whichvar = hvars[which(!is.na(h))]
   var_now  = get(whichvar)
   h        =h[!is.na(h)]
   h_vnam   = var_now$vnam [h]
   h_vmean  = paste0(h_vnam,"_mean")
   h_vqlwr  = paste0(h_vnam,"_qlwr")
   h_vqupr  = paste0(h_vnam,"_qupr")
   h_desc   = var_now$desc [h]
   h_short  = var_now$short[h]
   h_unit   = var_now$unit [h]
   h_legend = v == 1 


   # Temporary data table. We convert the classes back to factor.

   h_mmean$source = factor(h_mmean$source,levels=unique(h_mmean$source))

   # Initialise plot (we use line)
   gg_now = ggplot( data    = h_mmean
                  , mapping = aes_string( x      = "month"
                                        , group  = "source"
                                        , colour = "source"
                                        , fill   = "source"
                                        )
                  )#end ggplot
   gg_now = gg_now + scale_colour_manual(name="",aesthetics="colour",labels=leg_labels,values=leg_colours)
   gg_now = gg_now + scale_colour_manual(name="",aesthetics="fill"  ,labels=leg_labels,values=leg_colours)

   # We only add legend for a single plot.  Patchwork will fix this in the end.
   gg_now = gg_now + geom_ribbon( aes_string(ymin=h_vqlwr,ymax=h_vqupr)
                                , alpha       = alpha_ribbon
                                , show.legend = h_legend
                                , colour      = "transparent"
                                )#end geom_ribbon
   gg_now = gg_now + geom_line( aes_string(y=h_vmean)
                              , linewidth = 0.7
                              , show.legend = h_legend
                              )#end geom_line

   
   # Add local annotation
   gg_now = gg_now + labs(title=element_blank())
   gg_now = gg_now + scale_x_continuous( breaks = sequence(12)
                                       , labels = substring(month.abb,1,1)
                                       )#end scale_x_continuous
   gg_now = gg_now + xlab(element_blank())
   gg_now = gg_now + ylab(desc.unit(desc=h_short,unit=untab[[h_unit]],dxpr=TRUE))
   gg_now = gg_now + theme_bw( base_size      = gg_ptsz
                             , base_family    = "Helvetica"
                             , base_line_size = 0.5
                             , base_rect_size = 0.5
                               )#end theme_grey

   # Additional legend settings
   # if (h_legend) gg_now = gg_now + theme( legend.position = "bottom")

   # Axis settings
   gg_now = gg_now + theme( axis.text.x       = element_text( size   = gg_ptsz
                                                            , margin = unit(rep(0.35,times=4),"char")
                                                            )#end element_text
                          , axis.text.y       = element_text( size   = gg_ptsz
                                                            , margin = unit(rep(0.35,times=4),"char")
                                                            )#end element_text
                          , axis.ticks.length = unit(-0.2,"char")
                          , axis.title.y      = element_text( size = gg_ptsz * 0.9)
                          )#end theme

  # Write plot settings to the list.
  gg_emean[[h_vnam]] = gg_now
}#end for (v in seq_along(var_both)){

# Wrap plots then global settings and x axis.
gg_patch = wrap_plots(gg_emean)
gg_patch = gg_patch + guide_area() + plot_layout(guides="collect")
gg_patch = gg_patch + plot_annotation(tag_levels = "a")

# Save plots.
for (d in sequence(ndevice)){
   h_output = paste0("comparison_meanseason",case_now,".",gg_device[d])
   dummy    = ggsave( filename = h_output
                    , plot     = gg_patch
                    , device   = gg_device[d]
                    , path     = secy_path
                    , width    = gg_width
                    , height   = gg_height
                    , units    = gg_units
                    , dpi      = gg_depth
                    )#end ggsave

}#end for (d in sequence(n_device))
  
}


# If sought, plot images on screen
if (gg_screen) gg_patch


```




